### 2015-09-26 proj1.tex
Dummy LaTeX

\documentclass{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{amsmath}

\pagestyle{fancy}
\fancyhead{}
\fancyfoot{}
\rfoot{\thepage}
\lfoot{\includegraphics[height=20pt]{Logo}}

\usepackage{float}
\usepackage{listings, color, times, textcomp, float}
\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}
\lstset{language=Matlab, basicstyle=\scriptsize\ttfamily,breaklines=true,frame=single,morekeywords={matlab2tikz},
keywordstyle=\color{blue}, morekeywords=[2]{1}, keywordstyle=[2]{\color{black}}, identifierstyle=\color{black},
stringstyle=\color{mylilas}, commentstyle=\color{mygreen}, showstringspaces=false, numbers=left,
numberstyle={\tiny \color{black}}, numbersep=9pt, emph=[1]{for,end,break},emphstyle=[1]\color{red},
literate={~} {\texttildelow}{1}}

\title{MATH 3610 - Project \#1}
\author{Paul Chesnais (pmc85), Christopher Silvia (cps232), Ryan Vogan (rcv39)}
\date{\today}

\newcommand{\exo}[1]{\section*{Exercise #1}}
\newcommand{\prob}[1]{\section*{Problem #1}}
\newcommand{\quest}[1]{\section*{Question #1}}
\newcommand{\e}{&=}
\newcommand{\p}[1]{\times 10^{#1}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0.5pt}

\begin{document}
\maketitle
\thispagestyle{fancy}

\end{document}

### 2015-09-26 spread.m
Change makefile

@@ -7,16 +7,16 @@ vaccines = 4000;
 init_sick = 100;
 
 % Census data for age ranges
-num_children = ithaca_pop/10;
-num_teens = 2*ithaca_pop/5;
-num_adults = 2*ithaca_pop/5;
-num_seniors = ithaca_pop/10;
+num_children  = ithaca_pop/10;
+num_teens     = 2*ithaca_pop/5;
+num_adults    = 2*ithaca_pop/5;
+num_seniors   = ithaca_pop/10;
 
 % Values to pass to person classdef
-child = 1;
-teen = 2;
-adult = 3;
-senior = 4;
+child   = 1;
+teen    = 2;
+adult   = 3;
+senior  = 4;
 
 citizens = Person(0);
 
@@ -47,12 +47,13 @@ for i = 1:init_sick
   citizens(random_citizen).is_sick = true;
 end
 
-days = 20;
+months = 20;
 
-for d = 1:days
+for d = 1:months
   for citizen = citizens
     citizen.step();
   end
+
 end
 
 num_sick = 0;

### 2015-09-26 proj1.tex
moved matlab scripts to models directory
Added section on decision variables

@@
 
+\section{Decision Variables}
+
+The client has proposed two strategies:
+
+\begin{enumerate}
+\item Focus vaccination efforts on the ``most connected''.
+\item Focus vaccination efforts on the most frail or susceptible.
+\end{enumerate}
+
+In our models, we want to express our choice of policy as a ``decision variable''.
+Since we will be recieving 4000 vaccinations a month, our policy recommendation
+	will be how to distribute those 4000 vaccinations.
+If we identify $n$ categories of Ithacans, each month we will provide 
+	recommendations on which fraction of the 4000 vaccines to give to 
+	each category of Ithacans.
+
 \end{document}

### 2015-09-26 proj1.tex
Added section on evaluating models

@@
 
+We should note that each model won't necessarily divide the population into the
+	same number of categories.
+Each one will provide a different recommendation for the health officials.
+
+\section{Evaluation Models}
+
+Once we chose parameters for our model, we will run it and show its predictions.
+We expect that models should at least model the total number of sick
+	individuals.
+The model might also keep track of how many individuals die.
+We have thought of three different ways of evaluating our model:
+
+\begin{enumerate}
+\item Total Deaths
+\item Total Person-Days Spent Sick
+\item Time Until Disease is Eradicated
+\end{enumerate}
+
+Note that the models may only provide predictions for some of these criteria.
+
 \end{document}

### 2015-09-26 proj1.tex
adding disease spread with vaccine and with no vaccine

@@
 \usepackage{amsmath}
+\usepackage{tikz}
 
@@
 
-\section{Evaluation Models}
+\section{Evaluating Models}
 
@@
 
+\section{Simple Continuous Model}
+
+First, we propose a deterministic model which doesn't divide the population 
+	into sections at all.
+This model will help give us an idea of what the effect of dividing
+	the population into sections will be.
+In addition, it can give us an idea of how vaccination affects the spread of
+	disease within a population.
+
+\subsection{Disease Spread with No Vaccine}
+
+We include this derivation of the logistic model for disease spread
+	to compare with our derivation of the modified logistic model
+	for disease spread and vaccination.
+
+Suppose there are two groups of people, with $S$ representing the number of 
+	sick people, and $H$ represnting the number of healthy people.
+Suppose the probability of a single healthy person becoming sick is proportional
+	only to the total number of sick people.
+Since there are $H$ healthy people, the expected number of healthy people
+	who become sick should be proportional to $S H$.
+This model does not take into account recovery from sickness.
+If we introduce a parameter $r$, we can express this relationship as a system
+	of differential equations:
+
+\begin{align*}
+\frac{dH}{dt} & = - r S H \\
+\frac{dS}{dt} & = r S H \\
+\end{align*}
+
+Notice that the total number of people, $S + H$, is conserved, i.e.,
+
+\[ \frac{d}{dt} \left( S + H \right) = \frac{dS}{dt} + \frac{dH}{dt}
+	= r S H - r S H = 0 \]
+
+If we introduce another parameter, $K$, to represent the total population,
+	such that $S + H = K$ for all time, then the differential equations
+	can be decoupled:
+
+\begin{align*}
+\frac{dH}{dt} & = - r (K - H) H \\
+\frac{dS}{dt} & = r S (K - S) \\
+\end{align*}
+
+These are two independant logistic models.
+
+\subsection{Diease Spread with Vaccine}
+
+Suppose we introduce a third group, $V$, representing people who are vaccinated	
+	against the disease.
+Suppose we randomly vaccinate members of the healthy population,
+	a constant number each time step.
+This introduces a third parameter, $r_v$, the rate of vaccinations.
+The vaccinated population isn't seceptible to the disease.
+This process is modeled by the following system of diffential equations:
+
+\begin{align*}
+\frac{dS}{dt} & = r S H \\
+\frac{dH}{dt} & = - r S H - r_v \\
+\frac{dV}{dt} & = r_v
+\end{align*}
+
+This model is accurate only if all three of the populations remain
+	positive.
+This is important to remember.
+The number of healthy people will always be decreasing, so once
+	the number of healthy people hits zero, we stop the model.
+The sick people stay sick, and the vaccinated people stay vaccinated.
+
+The number of vaccinated people as a function of time is:
+
+\[ V(t) = r_v t \]
+
+Again, we find the conserved quantity $\left( S + H + V \right)$ and
+	name it $K$.
+For the model to be valid, we need $H = K - S - V \geq 0$.
+Therefore, the model is valid as long as:
+
+\[ K \geq S + r_v t \]
+
+We have now made these differential equations independant:
+
+\begin{align*}
+\frac{dS}{dt} & = r S \left( K - S - r_v t \right)\\
+\frac{dH}{dt} & = - r H \left( K - H - r_v t \right) - r_v \\
+\frac{dV}{dt} & = r_v t
+\end{align*}
+
 \end{document}

### 2015-09-26 models/spread.m
Just testing some stuff

@@ -12,34 +12,56 @@ num_teens     = 2*ithaca_pop/5;
 num_adults    = 2*ithaca_pop/5;
 num_seniors   = ithaca_pop/10;
 
-% Values to pass to person classdef
-child   = 1;
-teen    = 2;
-adult   = 3;
-senior  = 4;
+child = struct;
+child.comp_chance = 0.2;
+child.connectivity = 100;
+child.heal_chance = 0.3;
 
-citizens = Person(0);
+teen = struct;
+teen.comp_chance = 0.1;
+teen.connectivity = 200;
+teen.heal_chance = 0.2;
 
-% if ~(exist('citizens.mat','file') == 2)
+adult = struct;
+adult.comp_chance = 0.3;
+adult.connectivity = 150;
+adult.heal_chance = 0.1;
+
+senior = struct;
+senior.comp_chance = 0.7;
+senior.connectivity = 100;
+senior.heal_chance = 0.05;
+
+sick = zeros(1, ithaca_pop);
+vaccinated = zeros(1, ithaca_pop);
+healed = zeros(1, ithaca_pop);
+hospitalized = zeros(1, ithaca_pop);
+% connections = zeros(max([child.connectivity, teen.connectivity, adult.connectivity, senior.connectivity]), ithaca_pop);
+
 for i = 1:ithaca_pop
   if i <= num_children
-    citizens(i) = Person(child);
+    connections{i} = zeros(child.connectivity);
   elseif i <= (num_children + num_teens)
-    citizens(i) = Person(teen);
+    connections{i} = zeros(teen.connectivity);
   elseif i <= (num_children + num_teens + num_adults)
-    citizens(i) = Person(adult);
+    connections{i} = zeros(adult.connectivity);
   else
-    citizens(i) = Person(senior);
+    connections{i} = zeros(senior.connectivity);
   end
 end
 
-for i = citizens
-  while length(i.connections) < i.connectivity
+for i = 1:ithaca_pop
+  n = 1;
+  while n <= length(connections{i})
     random_citizen = round(rand*(ithaca_pop -1)) + 1;
-    if length(citizens(random_citizen)) < citizens(random_citizen).connectivity
-      citizens(random_citizen).connections(length(citizens(random_citizen))) = i.findobj;
-      i.connections(length(i.connections) + 1) = citizens(random_citizen).findobj;
+    if ~all(connections{random_citizen})
[…]

### 2015-09-27 proj1.tex
Closed form for S

@@
 \renewcommand{\footrulewidth}{0.5pt}
+
+\DeclareMathOperator\erf{erf}
 
@@
 
-We have now made these differential equations independant:
+\begin{tikzpicture}[scale=2.0]
+	\coordinate[label=below:{$0$}] (origin) at (0,0);
+	\draw[<->, thick] (0,2) node (saxis) [above] {$S$}
+	|- (3,0) node (taxis) [right] {$t$};
+
+	\coordinate[label=left:{$K$}] (sintercept) at (0,1.5);
+	\coordinate[label=below:{$\frac{K}{r_v}$}] (tintercept) at (2.2,0);
+	\draw (sintercept) -- (tintercept);
+	\fill[gray!20, opacity=0.8] (origin) -- (sintercept) -- (tintercept) -- cycle;
+
+	\draw[dotted] (sintercept) -- (3.0, 1.5);
+	\node (valid) at (0.7,0.4) {Valid Region};
+	\node (endregion) at (2.3, 1.0) {No-Change Region};
+\end{tikzpicture}
+
+In the above figure, once $S(t)$ hits the diagonal line,
+	it stays horizontal and keeps its value.
+
+We have now made these differential equations independant,
+	within the model's zone of validity:
 
@@
 
+We present a solution for $S(t)$:
+
+\begin{align}
+S(t) & = 
+	\frac{e^{ r t (K - \frac{ r_v t}{2})}}
+	{\frac{1}{S(0)} + \sqrt{\frac{\pi r}{2 r_v}}
+		e^{\frac12 \left( \frac{r}{r_v} \right) K^2}
+		\left(\erf\left(  \sqrt{\frac{r}{2 r_v}} (r_v t - K ) \right)
+			- \erf\left( - \sqrt{\frac{r}{2 r_v}} K \right) \right)}
+\end{align}
+
+From the above expression for the boundary of the valid region,
+	we get the following implicit expression for $t_f, S(t_f)$
+
+\begin{align}
+S(t_f) = 
+	\frac{e^{ r t_f (K - \frac{ r_v t_f }{2})}}
+	{\frac{1}{S(0)} + \sqrt{\frac{\pi r}{2 r_v}}
+		e^{\frac12 \left( \frac{r}{r_v} \right) K^2}
+		\left(\erf\left(  \sqrt{\frac{r}{2 r_v}} (r_v t_f - K ) \right)
+			- \erf\left( - \sqrt{\frac{r}{2 r_v}} K \right) \right)}
+	& = K - r_v t_f 
+\end{align}
+
 \end{document}

### 2015-09-27 models/Person.m
Trying out a different spread model

@@ -8,7 +8,7 @@ classdef Person < handle
     heal_chance       % Chance to heal after a given day
     comp_chance       % Chance of complications due to sickness
     connectivity      % Number of people exposed to in a given day
-    connections = Person.empty(0) % People connected to (sizeof(connections) == connectivity)
+    connections = Person.empty(0) % People connected to (length(connections) == connectivity)
   end
 
   methods

### 2015-09-27 models/spread.m

@@ -20,7 +20,6 @@ senior  = 4;
 
 citizens = Person(0);
 
-% if ~(exist('citizens.mat','file') == 2)
 for i = 1:ithaca_pop
   if i <= num_children
     citizens(i) = Person(child);
@@ -33,20 +32,30 @@ for i = 1:ithaca_pop
   end
 end
 
-for i = citizens
-  while length(i.connections) < i.connectivity
-    random_citizen = round(rand*(ithaca_pop -1)) + 1;
-    if length(citizens(random_citizen)) < citizens(random_citizen).connectivity
-      citizens(random_citizen).connections(length(citizens(random_citizen))) = i.findobj;
-      i.connections(length(i.connections) + 1) = citizens(random_citizen).findobj;
-    end
-  end
-end
+% if ~(exist('mapping.mat','file') == 2)
+%   for i = 1:length(citizens)
+%     mapping{i} = ones(1, citizens(i).connectivity);
+%     while length(citizens(i).connections) < citizens(i).connectivity
+%       random_citizen = round(rand*(ithaca_pop -1)) + 1;
+%       if length(citizens(random_citizen)) < citizens(random_citizen).connectivity
+%         citizens(random_citizen).connections(length(citizens(random_citizen))) = citizens(i).findobj;
+%         citizens(i).connections(length(citizens(i).connections) + 1) = citizens(random_citizen).findobj;
+%         mapping{i}(length(citizens(i).connections)) = random_citizen;
+%       end
+%     end
+%   end
 
-%   save 'citizens.mat' citizens;
+%   save 'mapping.mat' mapping;
 % else
-%   load 'citizens.mat';
+%   load 'mapping.mat';
+%   for i = 1:length(mapping)
+%     for n = 1:length(mapping{i})
+%       citizens(i).connections(n) = citizens(mapping{i}(n)).findobj;
+%     end
+%   end
 % end
+
+disp('Finished creating network.')
 
 % for i = 1:init_sick
 %   random_citizen = round(rand*(ithaca_pop -1)) + 1;

### 2015-09-27 proj1.tex

@@
 \usepackage{tikz}
+\usepackage{csvsimple}
+\usepackage{pgfplots}
 
@@
 
+Here is a family of values of $S(t)$.
+The total population and the rate of distributing vaccines are
+	held constant, while each curve corresponds to a different
+	value of the growth rate constant.
+Each curve starts with an initial condition at 5 percent of the
+	population, and the growth rate determines how many people
+	overall are infected.
+
+\paragraph{}
+\pgfplotstableread[col sep = comma]{data/s.csv}\mydata
+\begin{tikzpicture}
+	\begin{axis}[
+		legend pos = south east,
+		xmin = 0,
+		xmax = 6,
+		ymin = 0,
+		ymax = 1.3,
+		ylabel = {$S(t)$ (units of total population)},
+		xlabel = {$t$}
+	]
+	\foreach \index in {1, ..., 35} {%
+		\addplot[mark=""] table[x index = {0}, y index = {\index}]{\mydata};
+	}
+	\addplot[mark=""] {1 - x * 0.2};
+	\end{axis}
+\end{tikzpicture}
+\paragraph{}
+
 \end{document}

### 2015-09-27 models/Person.m
Hmm maybe the other way still works

@@ -34,22 +34,22 @@ classdef Person < handle
       end
     end
 
-    function step(self)
-      if self.is_sick
-        if rand < self.heal_chance
-          self.is_sick = false;
-          self.was_sick = true;
-        elseif rand < self.comp_chance
-          self.is_hospitalized = true;
-        else
-          for citizen = self.connections
-            if and(~citizen.is_vaccinated, and(~citizen.was_sick, ~citizen.is_sick))
-              citizen.is_sick = rand < 0.25; % randomly infect connection
-            end
-          end
-        end
-      end
-    end
+    % function step(self)
+    %   if self.is_sick
+    %     if rand < self.heal_chance
+    %       self.is_sick = false;
+    %       self.was_sick = true;
+    %     elseif rand < self.comp_chance
+    %       self.is_hospitalized = true;
+    %     else
+    %       for citizen = self.connections
+    %         if and(~citizen.is_vaccinated, and(~citizen.was_sick, ~citizen.is_sick))
+    %           citizen.is_sick = rand < 0.25; % randomly infect connection
+    %         end
+    %       end
+    %     end
+    %   end
+    % end
 
   end
 end

### 2015-09-27 models/spread.m

@@ -13,21 +13,25 @@ num_adults    = 2*ithaca_pop/5;
 num_seniors   = ithaca_pop/10;
 
 child = struct;
+child.age = 1;
 child.comp_chance = 0.2;
 child.connectivity = 100;
 child.heal_chance = 0.3;
 
 teen = struct;
+teen.age = 2;
 teen.comp_chance = 0.1;
 teen.connectivity = 200;
 teen.heal_chance = 0.2;
 
 adult = struct;
+adult.age = 3;
 adult.comp_chance = 0.3;
 adult.connectivity = 150;
 adult.heal_chance = 0.1;
 
 senior = struct;
+senior.age = 4;
 senior.comp_chance = 0.7;
 senior.connectivity = 100;
 senior.heal_chance = 0.05;
@@ -36,72 +40,54 @@ sick = zeros(1, ithaca_pop);
 vaccinated = zeros(1, ithaca_pop);
 healed = zeros(1, ithaca_pop);
 hospitalized = zeros(1, ithaca_pop);
-% connections = zeros(max([child.connectivity, teen.connectivity, adult.connectivity, senior.connectivity]), ithaca_pop);
+
+citizens = zeros(1, ithaca_pop);
 
 for i = 1:ithaca_pop
   if i <= num_children
-    connections{i} = zeros(child.connectivity);
+    citizens(i) = child;
   elseif i <= (num_children + num_teens)
-    connections{i} = zeros(teen.connectivity);
+    citizens(i) = teen;
   elseif i <= (num_children + num_teens + num_adults)
-    connections{i} = zeros(adult.connectivity);
+    citizens(i) = adult;
   else
-    connections{i} = zeros(senior.connectivity);
+    citizens(i) = senior;
   end
 end
 
-for i = 1:ithaca_pop
-  n = 1;
-  while n <= length(connections{i})
+for i = 1:init_sick
+  random_citizen = round(rand*(ithaca_pop -1)) + 1;
+  sick(random_citizen) = true;
+end
+
+months = 10;
+sick_per_month = zeros(months + 1, 4);
+
+for c = i:ithaca_pop
+  if sick(c)
+    sick_per_month(1, citizens(c).age) = sick_per_month(1, citizens(c).age) + 1;
+  end
+end
+
+for d = 1:months
+  vaccinated = 0;
+  while vaccinated ~= vaccines
     random_citizen = round(rand*(ithaca_pop -1)) + 1;
-    if ~all(connections{random_citizen})
[…]

### 2015-09-27 models/Person.m
Sigh...

@@ -8,7 +8,7 @@ classdef Person < handle
     heal_chance       % Chance to heal after a given day
     comp_chance       % Chance of complications due to sickness
     connectivity      % Number of people exposed to in a given day
-    connections = Person.empty(0) % People connected to (length(connections) == connectivity)
+    become_sick = false
   end
 
   methods
@@ -16,40 +16,24 @@ classdef Person < handle
       this.age = age;
       switch age
         case 1
-          this.comp_chance = 0.2;
-          this.connectivity = 100;
+          this.comp_chance = 0.1;
+          this.connectivity = 175;
           this.heal_chance = 0.3;
         case 2
-          this.comp_chance = 0.1;
+          this.comp_chance = 0.05;
           this.connectivity = 200;
           this.heal_chance = 0.2;
         case 3
-          this.comp_chance = 0.3;
+          this.comp_chance = 0.2;
           this.connectivity = 150;
           this.heal_chance = 0.1;
         case 4
-          this.comp_chance = 0.7;
-          this.connectivity = 100;
+          this.comp_chance = 0.4;
+          this.connectivity = 75;
           this.heal_chance = 0.05;
       end
+      % this.connectivity = this.connectivity * 2;
+      this.comp_chance = this.comp_chance / 10;
     end
-
-    function step(self)
-      if self.is_sick
-        if rand < self.heal_chance
-          self.is_sick = false;
-          self.was_sick = true;
-        elseif rand < self.comp_chance
-          self.is_hospitalized = true;
-        else
-          for citizen = self.connections
-            if and(~citizen.is_vaccinated, and(~citizen.was_sick, ~citizen.is_sick))
-              citizen.is_sick = rand < 0.25; % randomly infect connection
-            end
-          end
-        end
-      end
-    end
-
   end
 end

### 2015-09-27 models/spread.m

@@ -4,7 +4,8 @@ vaccines = 4000;
 ithaca_pop = ithaca_pop/10;
 vaccines = vaccines/10;
 
-init_sick = ithaca_pop/50;
+% init_sick = ithaca_pop/50;
+init_sick = 500;
 
 % Census data for age ranges
 num_children  = ithaca_pop/10;
@@ -32,64 +33,69 @@ for i = 1:ithaca_pop
   end
 end
 
-% if ~(exist('mapping.mat','file') == 2)
-%   for i = 1:length(citizens)
-%     mapping{i} = ones(1, citizens(i).connectivity);
-%     while length(citizens(i).connections) < citizens(i).connectivity
-%       random_citizen = round(rand*(ithaca_pop -1)) + 1;
-%       if length(citizens(random_citizen)) < citizens(random_citizen).connectivity
-%         citizens(random_citizen).connections(length(citizens(random_citizen))) = citizens(i).findobj;
-%         citizens(i).connections(length(citizens(i).connections) + 1) = citizens(random_citizen).findobj;
-%         mapping{i}(length(citizens(i).connections)) = random_citizen;
-%       end
-%     end
-%   end
+for c = citizens(round(rand(1,init_sick)*(ithaca_pop -1)) + 1)
+  c.is_sick = true;
+end
 
-%   save 'mapping.mat' mapping;
-% else
-%   load 'mapping.mat';
-%   for i = 1:length(mapping)
-%     for n = 1:length(mapping{i})
-%       citizens(i).connections(n) = citizens(mapping{i}(n)).findobj;
-%     end
-%   end
-% end
+deltas = 10;
+months = 10;
+sick_per_delta = zeros(1 + months*deltas, 5);
 
-disp('Finished creating network.')
+for c = citizens
+  if c.is_sick
+    sick_per_delta(1, c.age) = sick_per_delta(1, c.age) + 1;
+  end
+end
 
-% for i = 1:init_sick
-%   random_citizen = round(rand*(ithaca_pop -1)) + 1;
-%   citizens(random_citizen).is_sick = true;
-% end
+disp('Starting simulation');
 
-% months = 10;
-% sick_per_month = zeros(months + 1, 4);
+for d = 1:months*deltas
+  if ~mod(d, deltas)
+    disp(sprintf('Month: %d',d/deltas));
+  end
+  vaccinated = 0;
+  while vaccinated ~= (vaccines/deltas)
[…]

### 2015-09-27 models/citizen_type.m
What the hell...

@@ -1,0 +1,12 @@
+function citizen = citizen_type(n)
+  switch n
+    case 1
+      citizen = child;
+    case 2
+      citizen = teen;
+    case 3
+      citizen = adult;
+    case 4
+      citizen = senior;
+  end
+end

### 2015-09-27 models/spread.m

@@ -36,22 +36,22 @@ senior.comp_chance = 0.7;
 senior.connectivity = 100;
 senior.heal_chance = 0.05;
 
+citizens = zeros(1, ithaca_pop);
 sick = zeros(1, ithaca_pop);
 vaccinated = zeros(1, ithaca_pop);
 healed = zeros(1, ithaca_pop);
 hospitalized = zeros(1, ithaca_pop);
-
-citizens = zeros(1, ithaca_pop);
+become_sick = false(1, ithaca_pop);
 
 for i = 1:ithaca_pop
   if i <= num_children
-    citizens(i) = child;
+    citizens(i) = 1;
   elseif i <= (num_children + num_teens)
-    citizens(i) = teen;
+    citizens(i) = 2;
   elseif i <= (num_children + num_teens + num_adults)
-    citizens(i) = adult;
+    citizens(i) = 3;
   else
-    citizens(i) = senior;
+    citizens(i) = 4;
   end
 end
 
@@ -60,35 +60,64 @@ for i = 1:init_sick
   sick(random_citizen) = true;
 end
 
+deltas = 10;
 months = 10;
-sick_per_month = zeros(months + 1, 4);
+sick_per_delta = zeros(1 + months*deltas, 5);
 
 for c = i:ithaca_pop
   if sick(c)
-    sick_per_month(1, citizens(c).age) = sick_per_month(1, citizens(c).age) + 1;
+    sick_per_month(1, citizens(c)) = sick_per_month(1, citizens(c)) + 1;
   end
 end
+
+disp('starting sim')
 
 for d = 1:months
-  vaccinated = 0;
+  % if ~mod(d, deltas)
+    disp(sprintf('Month: %d',d/deltas));
+  % end
+  distributed = 0;
   while vaccinated ~= vaccines
     random_citizen = round(rand*(ithaca_pop -1)) + 1;
-    if ~citizens(random_citizen).is_sick
-      citizens(random_citizen).is_vaccinated = true;
-      vaccinated = vaccinated + 1;
+    if ~sick(random_citizen)
+      vaccinated(random_citizen) = true;
+      distributed = distributed + 1;
     end
   end
-  for citizen = citizens
-    % citizen.step();
-    random_citizens = rand()
+  for n = 1:ithaca_pop
+    t = citizen_type(citizens(n));
+    if sick(n)
+      if rand < (t.heal_chance/deltas)
+        sick(n) = false;
+        healed(n) = true;
[…]

### 2015-09-27 proj1.tex
Adding a log: work associated with parts of the tex file. Deterministic model for disease spread with vaccins

@@
 \rfoot{\thepage}
-\lfoot{\includegraphics[height=20pt]{Logo}}
 
@@
 
-\begin{tikzpicture}[scale=2.0]
-	\coordinate[label=below:{$0$}] (origin) at (0,0);
-	\draw[<->, thick] (0,2) node (saxis) [above] {$S$}
-	|- (3,0) node (taxis) [right] {$t$};
-
-	\coordinate[label=left:{$K$}] (sintercept) at (0,1.5);
-	\coordinate[label=below:{$\frac{K}{r_v}$}] (tintercept) at (2.2,0);
-	\draw (sintercept) -- (tintercept);
-	\fill[gray!20, opacity=0.8] (origin) -- (sintercept) -- (tintercept) -- cycle;
-
-	\draw[dotted] (sintercept) -- (3.0, 1.5);
-	\node (valid) at (0.7,0.4) {Valid Region};
-	\node (endregion) at (2.3, 1.0) {No-Change Region};
-\end{tikzpicture}
+\includegraphics[width=0.5\textwidth]{figures/admissible-region.pdf}
 
@@
 
-\paragraph{}
-\pgfplotstableread[col sep = comma]{data/s.csv}\mydata
-\begin{tikzpicture}
-	\begin{axis}[
-		legend pos = south east,
-		xmin = 0,
-		xmax = 6,
-		ymin = 0,
-		ymax = 1.3,
-		ylabel = {$S(t)$ (units of total population)},
-		xlabel = {$t$}
-	]
-	\foreach \index in {1, ..., 35} {%
-		\addplot[mark=""] table[x index = {0}, y index = {\index}]{\mydata};
-	}
-	\addplot[mark=""] {1 - x * 0.2};
-	\end{axis}
-\end{tikzpicture}
-\paragraph{}
+\includegraphics[width=0.5\textwidth]{figures/vaccination-model-curves-varying-s.pdf}
+
+\subsection{Pareto Optimizing Between Two Independent Populations}
+
+The model above modeled how one population responds to a combination
+	of disease spreading and vaccines.
+The model depends on $K$, the total population, $S(0)$, the initial
+	number of sick people, $r$, the rate of infection growth, 
+	and $r_v$, the vaccination rate.
+Of these parameters, in a real world setting, we can expect to know $K$
+	very precicely from census data, approximate $S(0)$ as small,
+	have very little idea about $r$, and control $r_v$.
+
+I want to extend this problem to two independant populations. 
+Each population will get different sets of constants,
+	$K_1$ and $K_2$, $S_1(0)$ and $S_2(0)$, $r_1$ and $r_2$.
+The goal will be to figure out how to optimally allocate the
+	vaccines between the two populations.
+The decision variable, $d$, will be the fraction of the total vaccination
+	rate I allocate to each population.
+Population 1 will be vaccinated at a rate $d r_v$, while population
+	2 will be vaccinated at a rate $(1 - d) r_v$.
+In this model, I try to minimize the final number of sick people in
+	each population.
+This will give me a pareto front.
+
+The following collection of pareto fronts comes from keeping the disease
+	growth rate the same for both, and varying its magnitude.
+
+\includegraphics[width=0.5\textwidth]{figures/vaccination-model-same-r-pareto-curves.pdf}
+
+This collection of pareto fronts comes from keeping the growth
+	rate for population 2 the same, and varying the growth
+	rate of population 1.
+
+\includegraphics[width=0.5\textwidth]{figures/vaccination-model-different-r-pareto-curves.pdf}
 

### 2015-09-27 models/Person.m
Looking good!!

@@ -16,24 +16,24 @@ classdef Person < handle
       this.age = age;
       switch age
         case 1
-          this.comp_chance = 0.1;
+          this.comp_chance = 0.05;
           this.connectivity = 175;
-          this.heal_chance = 0.3;
+          this.heal_chance = 0.5;
         case 2
           this.comp_chance = 0.05;
           this.connectivity = 200;
-          this.heal_chance = 0.2;
+          this.heal_chance = 0.4;
         case 3
-          this.comp_chance = 0.2;
+          this.comp_chance = 0.1;
           this.connectivity = 150;
-          this.heal_chance = 0.1;
+          this.heal_chance = 0.3;
         case 4
-          this.comp_chance = 0.4;
+          this.comp_chance = 0.6;
           this.connectivity = 75;
-          this.heal_chance = 0.05;
+          this.heal_chance = 0.2;
       end
       % this.connectivity = this.connectivity * 2;
-      this.comp_chance = this.comp_chance / 10;
+      % this.comp_chance = this.comp_chance / 10;
     end
   end
 end

### 2015-09-27 models/spread.m

@@ -38,7 +38,7 @@ for c = citizens(round(rand(1,init_sick)*(ithaca_pop -1)) + 1)
 end
 
 deltas = 10;
-months = 10;
+months = 3;
 sick_per_delta = zeros(1 + months*deltas, 5);
 
 for c = citizens
@@ -63,26 +63,26 @@ for d = 1:months*deltas
   end
   for citizen = citizens
     if citizen.is_sick
-      if rand < (citizen.heal_chance/deltas)
+      if (rand < (citizen.heal_chance/deltas))
         citizen.is_sick = false;
         citizen.was_sick = true;
-      elseif rand < (citizen.comp_chance/deltas)
+      elseif (rand < (citizen.comp_chance/deltas))
         citizen.is_sick = false;
         citizen.is_hospitalized = true;
       else
+        % random_citizens = normrnd();
         for c = citizens(round(rand(1,citizen.connectivity)*(ithaca_pop -1)) + 1)
-          % if ~c.is_vaccinated && ~c.was_sick && ~c.is_sick
-          c.become_sick = c.become_sick || 1;% rand < (1/deltas);
-          % end
+          if (~c.is_vaccinated && ~c.was_sick && ~c.is_sick)
+            c.become_sick = (c.become_sick || rand < (0.25/deltas));
+          end
         end
       end
     end
   end
   for c = citizens
-    if citizen.become_sick
-      disp('infect')
-      citizen.is_sick = true;
-      citizen.become_sick = false;
+    if c.become_sick
+      c.is_sick = true;
+      c.become_sick = false;
     end
   end
   for c = citizens
@@ -94,7 +94,7 @@ for d = 1:months*deltas
   end
 end
 
-disp('Plotting')
+disp('Plotting');
 
 f = figure('units', 'normalized', 'outerposition', [0 0 1 1]);
 bar(0:(1/deltas):months, sick_per_delta, 'stacked');

### 2015-09-28 final_submission/manuscript.tex
manuscript stub

\documentclass[titlepage]{article}

\usepackage{amsmath}

\title{Effective Vaccination Strategies for an H1N1 Epidemic in Ithaca, New York}
\author{Prepared by Paul Chesnais (pmc85), Christopher Silvia (cps232), \\ and Ryan Vogan (rcv39) in completion of Project 01 for MATH 3610}
\date{September 30, 2015}
\begin{document}
\maketitle

\section{Introduction}
	//TODO
\section{Model Assumptions}
	//TODO
\section{Parameter Values and Their Justifications}
	//TODO
\section{Simulation Algorithms}
	//TODO
\section{Model Predictions}
	//TODO
\section{Validation and Robustness Testing}
	//TODO
\section{Strengths and Weaknesses}
	//TODO
\section{Future Work}
	//TODO
\section{Conclusions}
	//TODO
\section{References}
	//TODO
\section{Appendix A}
	//TODO
\section{Appendix B}
	//TODO
\section{Individual Contributions}
	//TODO

\end{document}

### 2015-09-28 models/deterministic/basic_sir/sir.m
basic sir models

@@ -1,0 +1,16 @@
+clc; close all; clear;
+
+% ----- Model Parameters ----- %
+N = 60000;
+s0 = .99 * N;
+i0 = .01 * N;
+r0 = 0;
+beta = 1;
+gamma = .1;
+% ---------------------------- %
+
+t0 = 0;
+tf = 12;
+y0 = [s0,i0,r0];
+[T,Y] = ode45(@(t,y) sir_update(y, beta, gamma), [t0, tf], y0);
+plot(T,Y)

### 2015-09-28 models/deterministic/basic_sir/sir_with_vaccine_update.m

@@ -1,0 +1,12 @@
+function [y_prime] = sir_with_vaccine_update(y,beta,gamma,nu)
+s = y(1); 
+i = y(2);
+r = y(3);
+n = s + i + r;
+
+y_prime = zeros(3,1);
+y_prime(1) = -beta * s * i / n - nu * s;
+y_prime(2) = beta * s * i / n - gamma * i;
+y_prime(3) = gamma * i + nu * s;
+end
+

### 2015-09-29 models/Person.m
After the weekend

@@ -17,23 +17,17 @@ classdef Person < handle
       switch age
         case 1
           this.comp_chance = 0.05;
-          this.connectivity = 175;
+          this.connectivity = 200;
           this.heal_chance = 0.5;
         case 2
-          this.comp_chance = 0.05;
-          this.connectivity = 200;
-          this.heal_chance = 0.4;
-        case 3
           this.comp_chance = 0.1;
           this.connectivity = 150;
           this.heal_chance = 0.3;
-        case 4
+        case 3
           this.comp_chance = 0.6;
           this.connectivity = 75;
           this.heal_chance = 0.2;
       end
-      % this.connectivity = this.connectivity * 2;
-      % this.comp_chance = this.comp_chance / 10;
     end
   end
 end

### 2015-09-29 models/random_sample.m

@@ -1,0 +1,24 @@
+function people = random_sample(age)
+  switch age
+    case 1
+      ppl = normrnd(child_center, num_children/2);
+      ppl(ppl < 1) = 1;
+      ppl(ppl > ithaca_pop) = ithaca_pop;
+      people = ppl;
+    case 2
+      ppl = normrnd(teen_center, num_teens/2);
+      ppl(ppl < 1) = 1;
+      ppl(ppl > ithaca_pop) = ithaca_pop;
+      people = ppl;
+    case 3
+      ppl = normrnd(adult_center, num_adults/2);
+      ppl(ppl < 1) = 1;
+      ppl(ppl > ithaca_pop) = ithaca_pop;
+      people = ppl;
+    case 4
+      ppl = normrnd(senior_center, num_seniors/2);
+      ppl(ppl < 1) = 1;
+      ppl(ppl > ithaca_pop) = ithaca_pop;
+      people = ppl;
+  end
+end

### 2015-09-29 models/spread.m

@@ -1,101 +1,122 @@
-ithaca_pop = 60000;
-vaccines = 4000;
+function spread
+  ithaca_pop = 60000;
+  vaccines = 4000;
 
-ithaca_pop = ithaca_pop/10;
-vaccines = vaccines/10;
+  ithaca_pop = ithaca_pop/10;
+  vaccines = vaccines/10;
 
-% init_sick = ithaca_pop/50;
-init_sick = 500;
+  % init_sick = ithaca_pop/50;
+  init_sick = 500;
 
-% Census data for age ranges
-num_children  = ithaca_pop/10;
-num_teens     = 2*ithaca_pop/5;
-num_adults    = 2*ithaca_pop/5;
-num_seniors   = ithaca_pop/10;
+  % Census data for age ranges
+  num_juniors   = 0.6*ithaca_pop;
+  num_adults    = 0.3*ithaca_pop;
+  num_seniors   = 0.1*ithaca_pop;
 
-% Values to pass to person classdef
-child   = 1;
-teen    = 2;
-adult   = 3;
-senior  = 4;
+  % Values to pass to person classdef
+  junior_center  = num_juniors/2;
+  adult_center   = num_juniors + num_adults/2;
+  senior_center  = num_juniors + num_adults + num_seniors/2;
 
-citizens = Person(0);
+  citizens = Person(0);
 
-for i = 1:ithaca_pop
-  if i <= num_children
-    citizens(i) = Person(child);
-  elseif i <= (num_children + num_teens)
-    citizens(i) = Person(teen);
-  elseif i <= (num_children + num_teens + num_adults)
-    citizens(i) = Person(adult);
-  else
-    citizens(i) = Person(senior);
-  end
-end
-
-for c = citizens(round(rand(1,init_sick)*(ithaca_pop -1)) + 1)
-  c.is_sick = true;
-end
-
-deltas = 10;
-months = 3;
-sick_per_delta = zeros(1 + months*deltas, 5);
-
-for c = citizens
-  if c.is_sick
-    sick_per_delta(1, c.age) = sick_per_delta(1, c.age) + 1;
-  end
-end
-
-disp('Starting simulation');
-
-for d = 1:months*deltas
-  if ~mod(d, deltas)
-    disp(sprintf('Month: %d',d/deltas));
-  end
-  vaccinated = 0;
-  while vaccinated ~= (vaccines/deltas)
-    random_citizen = round(rand*(ithaca_pop -1)) + 1;
[…]

### 2015-09-29 proj1.tex
changed variable names

@@
 
-Suppose there are two groups of people, with $S$ representing the number of 
-	sick people, and $H$ represnting the number of healthy people.
+Suppose there are two groups of people, with $I$ representing the number of 
+	infected people, and $S$ represnting the number of people seceptible for the
+	disease.
 Suppose the probability of a single healthy person becoming sick is proportional
 	only to the total number of sick people.
-Since there are $H$ healthy people, the expected number of healthy people
-	who become sick should be proportional to $S H$.
+Since there are $S$ healthy people, the expected number of healthy people
+	who become sick should be proportional to $S I$.
 This model does not take into account recovery from sickness.
@@
 \begin{align*}
-\frac{dH}{dt} & = - r S H \\
-\frac{dS}{dt} & = r S H \\
+\frac{dI}{dt} & = - r S I \\
+\frac{dS}{dt} & = r S I \\
 \end{align*}
 
-Notice that the total number of people, $S + H$, is conserved, i.e.,
+Notice that the total number of people, $S + I$, is conserved, i.e.,
 
-\[ \frac{d}{dt} \left( S + H \right) = \frac{dS}{dt} + \frac{dH}{dt}
-	= r S H - r S H = 0 \]
+\[ \frac{d}{dt} \left( S + I \right) = \frac{dS}{dt} + \frac{dI}{dt}
+	= r S I - r S I = 0 \]
 
 If we introduce another parameter, $K$, to represent the total population,
-	such that $S + H = K$ for all time, then the differential equations
+	such that $S + I = K$ for all time, then the differential equations
 	can be decoupled:
@@
 \begin{align*}
-\frac{dH}{dt} & = - r (K - H) H \\
+\frac{dI}{dt} & = - r (K - I) I \\
 \frac{dS}{dt} & = r S (K - S) \\
@@
 
-These are two independant logistic models.
+The two differential equations shown here are now two independant, uncoupled
+	differential equations which we can solve seperately.
 
@@
 
-Suppose we introduce a third group, $V$, representing people who are vaccinated	
-	against the disease.
+Suppose we introduce a third group, $V$ for vaccinate, representing people who are 
+	immune from the diesease.
+Right now, the only way to become immune is through vaccination,
+	but we may add recovery in subsequent models.
 Suppose we randomly vaccinate members of the healthy population,
@@
 \begin{align*}
-\frac{dS}{dt} & = r S H \\
-\frac{dH}{dt} & = - r S H - r_v \\
+\frac{dI}{dt} & = r S I \\
+\frac{dS}{dt} & = - r S I - r_v \\
 \frac{dV}{dt} & = r_v
@@
 
-Again, we find the conserved quantity $\left( S + H + V \right)$ and
+Again, we find the conserved quantity $\left( S + I + V \right)$ and
 	name it $K$.
-For the model to be valid, we need $H = K - S - V \geq 0$.
+For the model to be valid, we need $S = K - I - V \geq 0$.
 Therefore, the model is valid as long as:
 
-\[ K \geq S + r_v t \]
+\[ K \geq I + r_v t \]
 
@@
 
-In the above figure, once $S(t)$ hits the diagonal line,
+In the above figure, once $I(t)$ hits the diagonal line,
 	it stays horizontal and keeps its value.
@@
 \begin{align*}
-\frac{dS}{dt} & = r S \left( K - S - r_v t \right)\\
-\frac{dH}{dt} & = - r H \left( K - H - r_v t \right) - r_v \\
+\frac{dI}{dt} & = r I \left( K - I - r_v t \right)\\
+\frac{dS}{dt} & = - r S \left( K - S - r_v t \right) - r_v \\
 \frac{dV}{dt} & = r_v t
@@
 
-We present a solution for $S(t)$:
+We present a solution for $I(t)$ \cite{walph}:
 
 \begin{align}
-S(t) & = 
+I(t) & = 
 	\frac{e^{ r t (K - \frac{ r_v t}{2})}}
@@
 From the above expression for the boundary of the valid region,
-	we get the following implicit expression for $t_f, S(t_f)$
+	we get the following implicit expression for $t_f, I(t_f)$
 
 \begin{align}
-S(t_f) = 
+I(t_f) = 
 	\frac{e^{ r t_f (K - \frac{ r_v t_f }{2})}}
@@
 
-Here is a family of values of $S(t)$.
+Here is a family of values of $I(t)$.
 The total population and the rate of distributing vaccines are
@@
 	of disease spreading and vaccines.
-The model depends on $K$, the total population, $S(0)$, the initial
+The model depends on $K$, the total population, $I(0)$, the initial
 	number of sick people, $r$, the rate of infection growth, 
@@
 Of these parameters, in a real world setting, we can expect to know $K$
-	very precicely from census data, approximate $S(0)$ as small,
+	very precicely from census data, approximate $I(0)$ as small,
 	have very little idea about $r$, and control $r_v$.
@@
 Each population will get different sets of constants,
-	$K_1$ and $K_2$, $S_1(0)$ and $S_2(0)$, $r_1$ and $r_2$.
+	$K_1$ and $K_2$, $I_1(0)$ and $I_2(0)$, $r_1$ and $r_2$.
 The goal will be to figure out how to optimally allocate the
@@
 
+\begin{thebibliography}{9}
+\bibitem{walph}
+	Wolfram Research Inc.,
+	\emph{WolframAlpha},
+	Retrieved September 26, 2015
+	http://www.wolframalpha.com/input/?i=%E2%88%AB+e^%28-a+x^2+%2B+b+x%29+dx
+	http://www.wolframalpha.com/input/?i=ds%2Fdt+%3D+r+s+%28K+-+s+-+a+t%29
+\end{thebibliography}
+
 \end{document}

### 2015-09-29 proj1.tex
added nontdimensionalizing

@@
 \begin{align*}
-\frac{dI}{dt} & = - r S I \\
-\frac{dS}{dt} & = r S I \\
+\frac{dS}{dt} & = - r S I \\
+\frac{dI}{dt} & = r S I \\
 \end{align*}
@@
 \begin{align*}
-\frac{dI}{dt} & = - r (K - I) I \\
-\frac{dS}{dt} & = r S (K - S) \\
+\frac{dS}{dt} & = - r S (K - S) \\
+\frac{dI}{dt} & = r (K - I) I \\
 \end{align*}
@@
 	differential equations which we can solve seperately.
+
+\subsubsection{Nondimensionalizing the Model}
+
+It would be helpful to make all of the constants dimensionless parameters.
+This will give us insight into the different scales that the model works at.
+
+To turn $I$ into a dimensionless parameter, we can divide by $K$.
+
+\begin{align*}
+\frac{d\left(\frac{I}{K}\right)}{dt} 
+	& = \left( r K \right) \left( \frac{I}{K} \right) \left( 1 - \frac{I}{K} \right)\\
+\frac{d\widetilde{I}}{dt} 
+	& = \widetilde{r} \widetilde{I} \left( 1 - \widetilde{I} \right) 
+\end{align*}
+
+This shows that the dynamics of $I$ do not depend on its size.
+For each value of the dimensionless constant $\widetilde{r}$,
+	we determine the behavior from it.
 

### 2015-09-29 models/deterministic/basic_sir/sir_01.m
Basic sir models with death

@@ -1,0 +1,26 @@
+clc; close all; clear;
+%====================%
+% Standard SIR model %
+%====================%
+
+% ----- Model Parameters ----- %
+s0 = .95;
+i0 = .05;
+r0 = .0;
+
+d = .25;
+nu = 1/d;
+R0 = 1.5;
+beta = R0 * nu;
+% ---------------------------- %
+
+t0 = 0;
+tf = 12;
+y0 = [s0;i0;r0];
+[T,Y] = ode45(@(t,y) sir_update_01(y, beta, nu), [t0, tf], y0);
+
+figure
+plot(T,Y)
+xlabel('Months since outbreak')
+ylabel('Fraction of population')
+legend('Susceptible', 'Infected', 'Removed')

### 2015-09-29 models/deterministic/basic_sir/sir_03.m

@@ -1,0 +1,34 @@
+clc; close all; clear;
+%============================================%
+% Standard SIR model with death and vaccines %
+%============================================%
+
+% ----- Model Parameters ----- %
+s0 = .95;
+i0 = .05;
+r0 = .0;
+d0 = .0;
+N = 60000;
+
+d = .25;
+nu = 1/d;
+R0 = 1.5;
+beta = R0 * nu;
+mu = 0.00001709; % * 20;
+gamma = 4000/N;
+% ---------------------------- %
+
+t0 = 0;
+tf = 12;
+y0 = [s0;i0;r0;d0];
+[T,Y] = ode45(@(t,y) sir_update_03(y, beta, nu, mu,gamma), [t0, tf], y0);
+
+figure
+plot(T,Y(:,1:3))
+xlabel('Months since outbreak') % x-axis label
+ylabel('Fraction of population') % y-axis label
+legend('Susceptible', 'Infected', 'Removed')
+
+epidemic_duration = T(find(Y(:,2) < .01 * max(Y(:,2)),1)) % months
+total_deaths = Y(end,4) * N                               % people
+total_suffering = trapz(T,Y(:,2)) * N                     % person-months

### 2015-09-29 models/deterministic/basic_sir/sir_update_03.m

@@ -1,0 +1,15 @@
+function [y_prime] = sir_update_03(y,beta,nu,mu,gamma)
+s = y(1); 
+i = y(2);
+if s < 0
+    gamma = 0;
+end
+y_prime = zeros(3,1);
+
+
+y_prime(1) = -beta * s * i - gamma;
+y_prime(2) = beta * s * i - nu * i - mu * i;
+y_prime(3) = nu * i;
+y_prime(4) = mu * i;
+end
+

### 2015-09-29 models/deterministic/basic_sir/sir_with_vaccine_update.m

@@ -1,12 +1,0 @@
-function [y_prime] = sir_with_vaccine_update(y,beta,gamma,nu)
-s = y(1); 
-i = y(2);
-r = y(3);
-n = s + i + r;
-
-y_prime = zeros(3,1);
-y_prime(1) = -beta * s * i / n - nu * s;
-y_prime(2) = beta * s * i / n - gamma * i;
-y_prime(3) = gamma * i + nu * s;
-end
-

### 2015-09-29 proj1.tex
modified

@@
 
+\subsection{Disease Spread between Coupled Populations}
+
+Consider three populations, with two connected only by an intermediate
+	one.
+
+\begin{align*}
+\frac{d\widetilde{I}_1}{dt} & = \left( \left( r_{11} K_1 \right) \widetilde{I}_1 
+			+ \left( r_{12} K_2 \right)  \widetilde{I}_2 \right)
+		\left( 1 - \widetilde{I}_1 - \left( \frac{\alpha_1}{K_1} \right) t \right)\\
+\frac{d\widetilde{I}_2}{dt} & = \left( \left( r_{22} K_2 \right) \widetilde{I}_2 
+			+ \left( r_{21} K_1 \right)  \widetilde{I}_1 
+			+ \left( r_{23} K_3 \right) \widetilde{I}_3 \right)
+		\left( 1 - \widetilde{I}_2 - \left( \frac{\alpha_2}{K_2} \right) t \right)\\
+\frac{d\widetilde{I}_3}{dt} & = \left( \left( r_{33} K_3 \right) \widetilde{I}_3 
+			+ \left( r_{32} K_2 \right)  \widetilde{I}_2 \right)
+		\left( 1 - \widetilde{I}_3 - \left( \frac{\alpha_3}{K_3} \right) t \right)
+\end{align*}
+
+If the second population is relatively small, then we might be able to 
+	decrease the total number of infections by focusing our entire efforts
+	on the smallest population to cut the link between the two populations.
+We then treat the vaccination problem as a pareto optimization problem, to distribute
+	the vaccines amongst the two populations.
+However this may not be better than splitting the vaccinations between the two.
+
 \begin{thebibliography}{9}
@@
 	Retrieved September 26, 2015
-	http://www.wolframalpha.com/input/?i=%E2%88%AB+e^%28-a+x^2+%2B+b+x%29+dx
-	http://www.wolframalpha.com/input/?i=ds%2Fdt+%3D+r+s+%28K+-+s+-+a+t%29
+	http://www.wolframalpha.com/input/?i=\%E2\%88\%AB+e\^\%28-a+x\^2+\%2B+b+x\%29+dx
+	http://www.wolframalpha.com/input/?i=ds\%2Fdt+\%3D+r+s+\%28K+-+s+-+a+t\%29
 \end{thebibliography}

### 2015-09-29 models/deterministic/coupled.m

@@ -1,0 +1,39 @@
+global r11 = 1.0;
+global r22 = 0.8;
+global r33 = 2.0;
+global r12 = 0.2;
+global r21 = 0.2;
+global r23 = 0.3;
+global r32 = 0.3;
+
+
+global ak1 = 0.4;
+global ak2 = 0.3;
+global ak3 = 0.1;
+
+function derivs = f( I, t)
+	global r11; global r12; global r21;
+	global r22; global r23; global r32;
+	global r33; global ak1; global ak2;
+	global ak3;
+
+	derivs = zeros(1,3);
+	derivs(1) = (r11 * I(1) + r12 * I(2)).*(1.0 - I(1) - ak1 * t);
+	derivs(2) = (r21 * I(1) + r22 * I(2) + r23 * I(3) ) .* (1.0 - I(2) - ak2 * t);
+	derivs(3) = (r23 * I(2) + r33 * I(3)) .* (1.0 - I(3) - ak3 * t);
+end
+
+T = linspace(0,6);
+X = lsode(@f, [0.02; 1e-8; 1e-8], T);
+i = 1;
+while (i < length(T)) && (X(i,1) + ak1 * T(i) < 1) && (X(i,2) + ak2 * T(i) < 1) && (X(i,3) + ak3 * T(i) < 1)
+	i= i + 1;
+end
+X(1:i,:)
+hold on
+plot(T(1:i), X(1:i,1), "color", "r", ";Ithaca;")
+plot(T(1:i), X(1:i,2), "color", "g", ";Faculty and Staff;")
+plot(T(1:i), X(1:i,3), "color", "b", ";Students;")
+legend("location", "northwest")
+hold off
+pause

### 2015-09-30 models/stochastic/Person.m
Weekly model

@@ -17,16 +17,16 @@ classdef Person < handle
       switch age
         case 1
           this.comp_chance = 0.05;
-          this.connectivity = 200;
-          this.heal_chance = 0.5;
+          this.connectivity = 100;
+          this.heal_chance = 0.95;
         case 2
           this.comp_chance = 0.1;
-          this.connectivity = 150;
-          this.heal_chance = 0.3;
+          this.connectivity = 50;
+          this.heal_chance = 0.8;
         case 3
-          this.comp_chance = 0.6;
-          this.connectivity = 75;
-          this.heal_chance = 0.2;
+          this.comp_chance = 0.75;
+          this.connectivity = 10;
+          this.heal_chance = 0.5;
       end
     end
   end

### 2015-09-30 models/stochastic/compute.m

@@ -1,0 +1,16 @@
+k = zeros(10, 2, 2);
+
+for n = 1:10
+  [p,q] = spread(1, 1);
+  k(n, 1, 1) = p;
+  k(n, 2, 1) = q;
+end
+
+for n = 1:10
+  [p,q] = spread(1, 1);
+  k(n, 1, 2) = p;
+  k(n, 2, 2) = q;
+end
+
+mean(k(:,:, 1))
+mean(k(:,:, 2))

### 2015-09-30 models/stochastic/random_vaccine.m

@@ -1,0 +1,40 @@
+function people = random_vaccine(age, len)
+  global ithaca_pop;
+  global vaccines;
+  global init_sick;
+  global num_juniors;
+  global num_adults;
+  global num_seniors;
+  global junior_center;
+  global adult_center;
+  global senior_center;
+
+  if nargin == 1
+    switch age
+      case 1
+        people = min(ithaca_pop, max(1, round(normrnd(junior_center, num_juniors/4))));
+      case 2
+        people = min(ithaca_pop, max(1, round(normrnd(adult_center, num_adults))));
+      case 3
+        people = min(ithaca_pop, max(1, round(normrnd(senior_center, num_seniors/2))));
+    end
+  else
+    switch age
+      case 1
+        ppl = round(normrnd(junior_center, num_juniors/4, 1, len));
+        ppl(ppl < 1) = 1;
+        ppl(ppl > ithaca_pop) = ithaca_pop;
+        people = ppl;
+      case 2
+        ppl = round(normrnd(adult_center, num_adults, 1, len));
+        ppl(ppl < 1) = 1;
+        ppl(ppl > ithaca_pop) = ithaca_pop;
+        people = ppl;
+      case 3
+        ppl = round(normrnd(senior_center, num_seniors/2, 1, len));
+        ppl(ppl < 1) = 1;
+        ppl(ppl > ithaca_pop) = ithaca_pop;
+        people = ppl;
+    end
+  end
+end

### 2015-09-30 models/stochastic/spread.m

@@ -1,44 +1,40 @@
-function spread
+function [peak, hosps] = spread(age_range, div, p)
+  if nargin == 1
+    div = 10;
+    p = false;
+  elseif nargin == 2
+    p = false;
+  end
+
+  global ithaca_pop;
+  global vaccines;
+  global init_sick;
+  global num_juniors;
+  global num_adults;
+  global num_seniors;
+  global junior_center;
+  global adult_center;
+  global senior_center;
+
   ithaca_pop = 60000;
-  vaccines = 4000;
+  vaccines = 4000/4; % Vaccines per week
 
-  ithaca_pop = ithaca_pop/10;
-  vaccines = vaccines/10;
+  ithaca_pop = ithaca_pop/div;
+  vaccines = vaccines/div;
 
-  % init_sick = ithaca_pop/50;
-  init_sick = 500;
+  init_sick = ithaca_pop/50;
 
   % Census data for age ranges
+  % num_babies    = 0.03*ithaca_pop;
   num_juniors   = 0.6*ithaca_pop;
   num_adults    = 0.3*ithaca_pop;
   num_seniors   = 0.1*ithaca_pop;
 
-  % Values to pass to person classdef
-  junior_center  = num_juniors/2;
-  adult_center   = num_juniors + num_adults/2;
-  senior_center  = num_juniors + num_adults + num_seniors/2;
+  junior_center   = num_juniors/4; % Center the norm closer to babies
+  adult_center    = num_juniors + num_adults/2;
+  senior_center   = num_juniors + num_adults + num_seniors/2;
 
   citizens = Person(0);
-
-  function people = random_sample(age)
-    switch age
-      case 1
-        ppl = round(normrnd(junior_center, num_juniors));
-        ppl(ppl < 1) = 1;
-        ppl(ppl > ithaca_pop) = ithaca_pop;
-        people = ppl;
-      case 2
-        ppl = round(normrnd(adult_center, num_adults));
-        ppl(ppl < 1) = 1;
-        ppl(ppl > ithaca_pop) = ithaca_pop;
-        people = ppl;
-      case 3
-        ppl = round(normrnd(senior_center, num_seniors));
-        ppl(ppl < 1) = 1;
-        ppl(ppl > ithaca_pop) = ithaca_pop;
-        people = ppl;
-    end
-  end
 
   for i = 1:ithaca_pop
     if i <= num_juniors
@@ -54,9 +50,9 @@ function spread
[…]

### 2015-09-30 models/stochastic/Person.m
Getting started on the writeup!

@@ -5,7 +5,7 @@ classdef Person < handle
     is_vaccinated = false
     is_hospitalized = false % If complications encur, move to hospital, stop infecting other people
     was_sick = false  % Check if citizen has already been sick, assuming one can only get sick of the same strain once
-    heal_chance       % Chance to heal after a given day
+    heal_chance = 1      % Chance to heal after a given day
     comp_chance       % Chance of complications due to sickness
     connectivity      % Number of people exposed to in a given day
     become_sick = false
@@ -18,15 +18,15 @@ classdef Person < handle
         case 1
           this.comp_chance = 0.05;
           this.connectivity = 100;
-          this.heal_chance = 0.95;
+          % this.heal_chance = 0.95;
         case 2
           this.comp_chance = 0.1;
           this.connectivity = 50;
-          this.heal_chance = 0.8;
+          % this.heal_chance = 0.8;
         case 3
           this.comp_chance = 0.75;
           this.connectivity = 10;
-          this.heal_chance = 0.5;
+          % this.heal_chance = 0.5;
       end
     end
   end

### 2015-09-30 models/stochastic/random_vaccine.m

@@ -21,7 +21,7 @@ function people = random_vaccine(age, len)
   else
     switch age
       case 1
-        ppl = round(normrnd(junior_center, num_juniors/4, 1, len));
+        ppl = round(normrnd(junior_center, num_juniors/2, 1, len));
         ppl(ppl < 1) = 1;
         ppl(ppl > ithaca_pop) = ithaca_pop;
         people = ppl;

### 2015-09-30 models/stochastic/spread.m

@@ -25,12 +25,11 @@ function [peak, hosps] = spread(age_range, div, p)
   init_sick = ithaca_pop/50;
 
   % Census data for age ranges
-  % num_babies    = 0.03*ithaca_pop;
   num_juniors   = 0.6*ithaca_pop;
   num_adults    = 0.3*ithaca_pop;
   num_seniors   = 0.1*ithaca_pop;
 
-  junior_center   = num_juniors/4; % Center the norm closer to babies
+  junior_center   = num_juniors/2;
   adult_center    = num_juniors + num_adults/2;
   senior_center   = num_juniors + num_adults + num_seniors/2;
 

### 2015-09-30 models/deterministic/numerical_sir/sir_01.m
done sir models

@@ -1,7 +1,7 @@
 clc; close all; clear;
-%====================%
-% Standard SIR model %
-%====================%
+%===========%
+% SIR model %
+%===========%
 
 % ----- Model Parameters ----- %
 s0 = .95;
@@ -15,7 +15,7 @@ beta = R0 * nu;
 % ---------------------------- %
 
 t0 = 0;
-tf = 12;
+tf = 8;
 y0 = [s0;i0;r0];
 [T,Y] = ode45(@(t,y) sir_update_01(y, beta, nu), [t0, tf], y0);
 

### 2015-09-30 models/deterministic/numerical_sir/sir_update_01.m

@@ -3,7 +3,7 @@ s = y(1);
 i = y(2);
 y_prime = zeros(3,1);
 y_prime(1) = -beta * s * i;
-y_prime(2) = beta * s * i - nu * i;
+y_prime(2) = i * (beta * s - nu);
 y_prime(3) = nu * i;
 end
 

### 2015-09-30 models/deterministic/numerical_sir/sir_update_02.m

@@ -3,7 +3,7 @@ s = y(1);
 i = y(2);
 y_prime = zeros(3,1);
 y_prime(1) = -beta * s * i;
-y_prime(2) = beta * s * i - nu * i - mu * i;
+y_prime(2) = i * (beta * s - nu - mu);
 y_prime(3) = nu * i;
 y_prime(4) = mu * i;
 end

### 2015-09-30 models/deterministic/numerical_sir/sir_update_04.m

@@ -1,0 +1,14 @@
+function [y_prime] = sir_update_04(y,beta,nu,mu,gamma)
+s = y(1:4);
+i = y(5:8);
+gamma(s<=0) = 0;
+
+
+y_prime = zeros(16,1);
+for k = 1:4
+    y_prime(k) = -beta(k,:) * i * s(k) - gamma(k);
+    y_prime(k+4) = beta(k,:) * i * s(k) - i(k) * (nu + mu(k));
+    y_prime(k+8) = nu * i(k) + gamma(k);
+    y_prime(k+12) = mu(k) * i(k);
+end
+

### 2015-09-30 proj1.tex

--- 2015-09-29 proj1.tex
@@
 \usepackage{float}
-\usepackage{listings, color, times, textcomp, float}
+\usepackage{listings, color, times, textcomp, float, hyperref, subfigure}
 \definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
@@
 
+\section{Stochastic Model}
+\subsection{A naive model}
+\par At first, we made a simple, arbitrary stochastic model just to get a rough idea of how the disease would spread.
+Over the course of 5 months, we would look at the spread of the disease given certain conditions.
+Every citizen fell into an age category, according to demographic figures from \emph{areaConnect} \cite{areconnect}.
+Then, to model the spread, we associated arbitrary recovery rates, complication rates and number of people talked to per month to each age group, giving us the following table:
+\begin{figure}[H]
+\centering
+\begin{tabular}{c|c|c|c|c}
+Age group & Heal rate & Complication chance & Connectivity & Population percentage\\ \hline
+Kids & 0.5 & 0.2 & 20 & 10\%\\
+Teens & 0.4 & 0.1 & 30 & 40\%\\
+Adults & 0.3 & 0.3 & 25 & 40\%\\
+Seniors & 0.2 & 0.7 & 10 & 10\%
+\end{tabular}
+\end{figure}
+
+\par The Heal rate and Complication chance columns represent the odds of a given citizen to recover from the disease in a month.
+The Connectivity column represents the number of people each sick citizen has a chance to infect, uniformly sampled from the entire population.
+We would then infect 500 people initially, and let the simulation run its course.
+In this particular example, we vaccinate randomly.
+We also assume that a person who has been sick cannot be sick again, and that when complications arise in a given patient, they are taken to the hospital and do not keep infecting the rest of the population.
+Here's how the spread of disease looks:
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{figures/Naive-model.png}
+\end{figure}
+The parts of each bars represent the respective number of citizens in a given age group that are sick. The black bars on the bottom represent how many people have been hospitalized so far.
+\par Unfortunately, research and according to a source from the CIC \cite{CIC} exposed many flaws in this model.
+Mostly the fact that the odds of a given citizen staying sick for more than a month are extremely low, unlike in this model where they are actually quite high.
+In fact, every age range stays sick for just about the same amount of time, with seniors sick for 2 weeks.
+The CIC data also exposed something very interesting, which is that seniors are not the ones in the most danger from the flu, kids smaller than the age of 5 are.
+This gave us reason to use a model that looks at the spread of the disease on a weekly basis.
+
+\subsection{A more justified approach}
+\par As mentioned in earlier, the CIC data gave us a lot to think about in terms of refining this model.
+First of all, we looked at the effects of the population of children in Ithaca, where children are citizens younger than 5 years of age.
+Since their risks of complications are so much higher, a good idea would be to immediately vaccinate all of them to avoid a devastating death toll.
+Fortunately, according to \emph{areaConnect}, there are less than 2,000 children in Ithaca, meaning that they can all be vaccinated as soon as the shipment of vaccines arrives.
+This also means that we can cut out an entire segment of the population, and generalize citizens younger than 25 years of age to a single slice.
+\par We also have more concrete data on how long citizens stay sick, and the likelihood of complications.
+Given that all data is given in terms of weeks, we change the model to compute the spread on a weekly basis.
+For simplicity and rounding, we assume that there are 4 weeks in a month, and that all citizens, regardless of age, have a 90\% of healing every week.
+According to the CIC, a patient infected by the H1N1 virus is only contagious when showing symptoms.
+Therefore a citizen only infects other citizens when they are sick.
+Given what we know about the H1N1 virus and how contagious it is, we take 80\% to be the chance that a sick citizen infects a healthy one, with the 20\% accounting for sanitary precautions like hand sanitizer.
+\par
+
+\begin{figure}[H]
+\centering
+\begin{tabular}{c|c|c|c|c}
+Age group & Complication chance & Connectivity & Population percentage\\ \hline
+Junior & 0.05 & 100 & 100\%\\
+Adults & 0.1 & 50 & 50\%\\
+Seniors & 0.75 & 25 & 10\%
+\end{tabular}
+\end{figure}
+
 \begin{thebibliography}{9}
+\bibitem{areconnect}
+	\url{http://ithaca.areaconnect.com/statistics.htm}
+\bibitem{CIC}
+	\url{http://www.cdc.gov/h1n1flu/surveillanceqa.htm}
 \bibitem{walph}
@@
 	Retrieved September 26, 2015
-	http://www.wolframalpha.com/input/?i=%E2%88%AB+e^%28-a+x^2+%2B+b+x%29+dx
-	http://www.wolframalpha.com/input/?i=ds%2Fdt+%3D+r+s+%28K+-+s+-+a+t%29
+	\url{http://www.wolframalpha.com/input/?i=ds%2Fdt+%3D+r+s+%28K+-+s+-+a+t%29}
+	\url{http://www.wolframalpha.com/input/?i=%E2%88%AB+e^%28-a+x^2+%2B+b+x%29+dx}
 \end{thebibliography}

### 2015-09-30 models/stochastic/compute.m
No more deltas

@@ -1,13 +1,13 @@
 k = zeros(10, 2, 2);
 
 for n = 1:10
-  [p,q] = spread(1, 1);
+  [p,q] = spread(1, 10);
   k(n, 1, 1) = p;
   k(n, 2, 1) = q;
 end
 
 for n = 1:10
-  [p,q] = spread(1, 1);
+  [p,q] = spread(3, 10);
   k(n, 1, 2) = p;
   k(n, 2, 2) = q;
 end

### 2015-09-30 models/stochastic/kvalues.m

@@ -1,0 +1,23 @@
+k = zeros(10, 2, 2);
+
+k(:,:,1) =[59.23, 6188
+61.06, 6097
+60.64, 6141
+59.66, 6203
+59.84, 6194
+60.72, 6212
+60.35, 6210
+60.36, 6411
+58.65, 6006
+60.07, 6232]
+
+k(:,:,2)=[59.34, 6240
+60.28, 6242
+59.75, 6182
+59.71, 5974
+59.85, 6156
+60.5 , 6200
+60.16, 6146
+60.28, 6197
+60.32, 6159
+59.98, 6240]

### 2015-09-30 models/stochastic/spread.m

@@ -1,4 +1,4 @@
-function [peak, hosps] = spread(age_range, div, p)
+function [peak, hosps, sick_per_delta] = spread(age_range, div, p)
   if nargin == 1
     div = 10;
     p = false;
@@ -22,7 +22,7 @@ function [peak, hosps] = spread(age_range, div, p)
   ithaca_pop = ithaca_pop/div;
   vaccines = vaccines/div;
 
-  init_sick = ithaca_pop/50;
+  init_sick = ithaca_pop/20;
 
   % Census data for age ranges
   num_juniors   = 0.6*ithaca_pop;
@@ -33,74 +33,118 @@ function [peak, hosps] = spread(age_range, div, p)
   adult_center    = num_juniors + num_adults/2;
   senior_center   = num_juniors + num_adults + num_seniors/2;
 
-  citizens = Person(0);
+  heal_chance = 0.8;
 
-  for i = 1:ithaca_pop
-    if i <= num_juniors
-      citizens(i) = Person(1);
-    elseif i <= (num_juniors + num_adults)
-      citizens(i) = Person(2);
-    else
-      citizens(i) = Person(3);
+  age = zeros(1, ithaca_pop);
+  sick = zeros(1, ithaca_pop);
+  hospitalized = zeros(1, ithaca_pop);
+  vaccinated = zeros(1, ithaca_pop);
+
+  function conn = connectivity(age)
+    switch age
+      case 1
+        conn = 1000/div;
+      case 2
+        conn = 30/div;
+      case 3
+        conn = 10/div;
     end
   end
 
-  for c = citizens(round(rand(1,init_sick)*(ithaca_pop - 1)) + 1)
-    c.is_sick = true;
+  function comp = comp_chance(age)
+    switch age
+      case 1
+        comp = 0.05;
+      case 2
+        comp = 0.2;
+      case 3
+        comp = 0.75;
+    end
+  end
+
+  for i = 1:ithaca_pop
+    if i <= num_juniors
+      age(i) = 1;
+    elseif i <= (num_juniors + num_adults)
+      age(i) = 2;
+    else
+      age(i) = 3;
+    end
   end
 
   deltas = 5;
   weeks = 6; % 3 months
   sick_per_delta = zeros(1 + weeks*deltas, 4);
 
-  for c = citizens
-    if c.is_sick
[…]

### 2015-09-30 models/stochastic/compute.m
Good stuff, good stuff. Totally uncorrelated though

@@ -1,13 +1,13 @@
 k = zeros(10, 2, 2);
 
 for n = 1:10
-  [p,q] = spread(1, 10);
+  [p,q] = spread(1);
   k(n, 1, 1) = p;
   k(n, 2, 1) = q;
 end
 
 for n = 1:10
-  [p,q] = spread(3, 10);
+  [p,q] = spread(3);
   k(n, 1, 2) = p;
   k(n, 2, 2) = q;
 end

### 2015-09-30 models/stochastic/spread.m

@@ -33,17 +33,25 @@ function [peak, hosps, sick_per_delta] = spread(age_range, div, p)
   adult_center    = num_juniors + num_adults/2;
   senior_center   = num_juniors + num_adults + num_seniors/2;
 
-  heal_chance = 0.8;
+  heal_chance = 0.5;
 
   age = zeros(1, ithaca_pop);
   sick = zeros(1, ithaca_pop);
   hospitalized = zeros(1, ithaca_pop);
   vaccinated = zeros(1, ithaca_pop);
 
+  function v = ifelse(c, t, f)
+    if c
+      v = t;
+    else
+      v = f;
+    end
+  end
+
   function conn = connectivity(age)
     switch age
       case 1
-        conn = 1000/div;
+        conn = 50/div;
       case 2
         conn = 30/div;
       case 3
@@ -60,6 +68,7 @@ function [peak, hosps, sick_per_delta] = spread(age_range, div, p)
       case 3
         comp = 0.75;
     end
+    % comp = 0.95;
   end
 
   for i = 1:ithaca_pop
@@ -72,9 +81,8 @@ function [peak, hosps, sick_per_delta] = spread(age_range, div, p)
     end
   end
 
-  deltas = 5;
-  weeks = 6; % 3 months
-  sick_per_delta = zeros(1 + weeks*deltas, 4);
+  weeks = 12; % 3 months
+  sick_per_delta = zeros(1 + weeks, 4);
 
   for c = round(rand(1,init_sick)*(ithaca_pop - 1)) + 1
     sick(c) = true;
@@ -84,72 +92,48 @@ function [peak, hosps, sick_per_delta] = spread(age_range, div, p)
   n = 1;
   for w = 1:weeks
     disp(sprintf('Week: %d', w));
-    for d = 1:deltas
-      vax = 0;
-      peep = 1;
-      switch age_range
-        case 1
-          while vax ~= (vaccines/deltas)
-            if ~vaccinated(peep)
-              vaccinated(peep) = true;
-              vax = vax + 1;
-            end
-            peep = peep + 1;
-            if peep == ithaca_pop
-              break
-            end
-          end
-        case 3
-          while vax ~= (vaccines/deltas)
-            if ~vaccinated(ithaca_pop - peep + 1)
-              vaccinated(ithaca_pop - peep + 1) = true;
-              vax = vax + 1;
-            end
-            peep = peep + 1;
-            if peep == ithaca_pop
-              break
-            end
-          end
[…]

### 2015-09-30 final_submission/manuscript.tex
updated manuscript template

@@
 \section{Introduction}
+	For this project, we were tasked with modeling a H1N1 outbreak in Ithaca, New York.
+The goal of this effort was to provide local government officials with a recommended plan for distributing 4000 vaccines among the city's approximately 60000 inhabitants.
+The strategies that officials are particularly interested in are:
+	\begin{itemize}
+		\item[1.]
+			Focus vaccination efforts on more connected individuals
+		\item[2.]
+			Focus vaccination efforts on more frail or susceptible individuals
+	\end{itemize}
+	They wanted to know which strategy would be better under two possible scenarios:
+	\begin{itemize}
+		\item[1.]
+			An outbreak of the H1N1 virus
+		\item[2.]
+			An outbreak of a new H1N1 strain that is twenty times as deadly
+	\end{itemize}
+	When evaluating possible strategies, we considered the following two metrics:
+	\begin{itemize}
+		\item[1.]
+			Total number of deaths
+		\item[2.]
+			Total number of person-months of infection
+	\end{itemize}
+	Therefore, a particular distribution of vaccines is considered better if it saves more lives and prevents more suffering.
+
+\section{Deterministic Model}
+\subsection{Algorithm}
+
+\subsection{Assumptions and Parameter Justifications}
 	//TODO
-\section{Model Assumptions}
+\subsection{Predictions}
 	//TODO
-\section{Parameter Values and Their Justifications}
+\subsection{Validation and Robustness Testing}
 	//TODO
-\section{Simulation Algorithms}
+\subsection{Strengths and Weaknesses}
 	//TODO
-\section{Model Predictions}
+\subsection{Future Work}
 	//TODO
-\section{Validation and Robustness Testing}
+
+\section{Stochastic Model}
+\subsection{Algorithm}
 	//TODO
-\section{Strengths and Weaknesses}
+\subsection{Assumptions and Parameter Justifications}
 	//TODO
-\section{Future Work}
+\subsection{Predictions}
 	//TODO
+\subsection{Validation and Robustness Testing}
+	//TODO
+\subsection{Strengths and Weaknesses}
+	//TODO
+\subsection{Future Work}
+	//TODO
+
 \section{Conclusions}
@@
 	//TODO
-\section{Appendix A}
-	//TODO
-\section{Appendix B}
+\section{Source Code}
 	//TODO

### 2015-09-30 final_submission/manuscript.tex
SIR algorithm and assumptions done

@@
 \section{Deterministic Model}
-\subsection{Algorithm}
+\subsection{Algorithm and Assumptions}
+	Our initial research on epidemiology discovered a common model known as the SIR approach \cite{stanford}.
+In this model, the population is segmented into three compartments that represent susceptible, infected, and removed states.
+This model assumes:
+	\begin{itemize}
+		\item[1.]
+			A constant population size $N$
+		\item[2.]
+			Constant rates of transmission and recovery
+		\item[3.]
+			No births or deaths
+		\item[4.]
+			A well-mixed population
+	\end{itemize}
+	Define $S$ to be the number of susceptible individuals, $I$ to be the number of infected individuals, $R$ to be the number of recovered individuals, and $N$ to be the total population of Ithaca.
+Now normalize these variables  to consider the fraction of the total population in each compartment, such that $s = S/N$, $i = I/N$, $r = R/N$.
+Then the model is governed by the following three ordinary differential equations \cite{stanford}:
+	\[
+		\frac{ds}{dt} = - \beta s i
+	\]
+	\[
+		\frac{di}{dt} = \beta s i - \gamma i
+	\]
+	\[
+		\frac{dr}{dt} = \gamma i
+	\]
+	Here $\beta$ represents the effective contact rate, and $\gamma$ is the recovery rate \cite{stanford}.
+It is important to note that in this variant of the SIR model, we interpreted R to stand for Recovered rather than removed.
+Traditionally, the removed state represents all individuals who are not infected and are incapable of becoming infected again.
+This encompasses both individuals who recovered and those who died.
+We chose to seperate mortality rates out of the removal rate so that this parameter could be adjusted for the deadlier virus strain.
+Then the system of equations becomes:
+	\[
+		\frac{ds}{dt} = - \beta s i
+	\]
+	\[
+		\frac{di}{dt} =  i (\beta s -\gamma - \mu)
+	\]
+	\[
+		\frac{dr}{dt} = \gamma i
+	\]
+	\[
+		\frac{d\delta}{dt} = \mu i
+	\]
+	Here, $\mu$ is the death rate, and $\delta$ is the fraction of the population that has died from the swine flu.
+Therefore, this adds a fourth compartment, where a sick person can either recover or die.
+The final parameter necessary to model vaccination strategies on a homogeneous population is the vaccination rate itself.
+When this is added, the system becomes:
+	\[
+		\frac{ds}{dt} = - \beta s i - \nu
+	\]
+	\[
+		\frac{di}{dt} =  i (\beta s -\gamma - \mu)
+	\]
+	\[
+		\frac{dr}{dt} = \gamma i + \nu
+	\]
+	\[
+		\frac{d\delta}{dt} = \mu i
+	\]
+	Where $\nu$ is the fraction of the total population that gets vaccinated in a single month.
+This does not depend on the fractions of susceptible or infected individuals because we have a constant $4000$ vaccines per month.
+In the actual implementation of this model, it was necessary to code in a conditional statement that set this constant to $0$ once the susceptible population completely diminished, otherewise this would result in negative population values for that compartment.
+Also, note that this assumes that the number of vaccines distributed grows linearly throughout the month, finally reaching $4000$ at the very end.\\
 
-\subsection{Assumptions and Parameter Justifications}
+	The final missing peice of this model was to do away with the well-mixed assumption.
+This was crucial for us to make the distinction between highly-connected and frail individuals as requested by the government officials.
+We decided to segment the population into four categories.
+They are babies/toddlers ($<$ 4 years old), children/teens (between 5 and 24 years old), adults (between 25 and 65 years old), and seniors ($>$ 65 years old).
+These categorizations were chosen based on evidence for mortality rates that will be discussed in the ``Parameter Justifications'' section.
+This effectively quadruples our number of compartments, where the final set of ODE's is then:
+	\[
+		\frac{ds_k}{dt} = - (\beta_{k,1} i_1 + \beta_{k,2} i_2 + \beta_{k,3} i_3 + \beta_{k,4} i_4) s_k - \nu_k
+	\]
+	\[
+		\frac{di_k}{dt} =  (\beta_{k,1} i_1 + \beta_{k,2} i_2 + \beta_{k,3} i_3 + \beta_{k,4} i_4) s_k - i_k (\gamma + \mu_k)
+	\]
+	\[
+		\frac{dr_k}{dt} = \gamma i_k + \nu_k
+	\]
+	\[
+		\frac{d\delta_k}{dt} = \mu_k i_k
+	\]
+	Here, the subscript k indicates that the parameter belongs to the kth segment of the population, where k = 1 for babies/toddlers, k = 2 for children/teens, k=3 for adults, and k=4 for seniors.
+$\beta_{i,j}$ represents the rate of contact of an individual of type $i$ from an individual of type $j$.
+The values of these parameters come from a matrix called the ?
+that will be discussed in the ``Parameter Justifications'' section.
+Finally, note that there is no subscript on the $\gamma$ parameter, meaning we assume the recovery rate to be constant for all four age group.
+This is because recovery rate is related to the duration of an infection, as will be discussed in ``Parameter Justifications''.\\
+
+	We used MATLAB's built in ode45 solver on this system of equations to compute the fractions of suscpetible, infected, recovered, and dead individuals for each sub-segment of the population over time.
+\subsection{Parameter Justifications}
 	//TODO
@@
 \section{Stochastic Model}
-\subsection{Algorithm}
+\subsection{Algorithm and Assumptions}
 	//TODO
-\subsection{Assumptions and Parameter Justifications}
+\subsection{Parameter Justifications}
 	//TODO
@@
 \section{References}
-	//TODO
+	\begin{thebibliography}{9}
+	\bibitem{areconnect}
+		\url{http://ithaca.areaconnect.com/statistics.htm}
+	\bibitem{CIC}
+		\url{http://www.cdc.gov/h1n1flu/surveillanceqa.htm}
+	\bibitem{walph}
+		Wolfram Research Inc.,
+		\emph{WolframAlpha},
+		Retrieved September 26, 2015
+		\url{http://www.wolframalpha.com/input/?i=ds%2Fdt+%3D+r+s+%28K+-+s+-+a+t%29}
+		\url{http://www.wolframalpha.com/input/?i=%E2%88%AB+e^%28-a+x^2+%2B+b+x%29+dx}
+	\bibitem{stanford}
+		\url{http://web.stanford.edu/~jhj1/teachingdocs/Jones-on-R0.pdf}
+	\end{thebibliography}
+
 \section{Source Code}

### 2015-09-30 models/deterministic/numerical_sir/sir_update_04.m

@@ -1,14 +1,14 @@
-function [y_prime] = sir_update_04(y,beta,nu,mu,gamma)
+function [y_prime] = sir_update_04(y,beta,gamma,mu,nu)
 s = y(1:4);
 i = y(5:8);
-gamma(s<=0) = 0;
+nu(s<=0) = 0;
 
 
 y_prime = zeros(16,1);
 for k = 1:4
-    y_prime(k) = -beta(k,:) * i * s(k) - gamma(k);
-    y_prime(k+4) = beta(k,:) * i * s(k) - i(k) * (nu + mu(k));
-    y_prime(k+8) = nu * i(k) + gamma(k);
+    y_prime(k) = -beta(k,:) * i * s(k) - nu(k);
+    y_prime(k+4) = beta(k,:) * i * s(k) - i(k) * (gamma + mu(k));
+    y_prime(k+8) = gamma * i(k) + nu(k);
     y_prime(k+12) = mu(k) * i(k);
 end
 

### 2015-09-30 proj1.tex

@@
 
+\subsection{Disease Spread between Coupled Populations}
+
+Consider three populations, with two connected only by an intermediate
+	one.
+
+\begin{align*}
+\frac{d\widetilde{I}_1}{dt} & = \left( \left( r_{11} K_1 \right) \widetilde{I}_1 
+			+ \left( r_{12} K_2 \right)  \widetilde{I}_2 \right)
+		\left( 1 - \widetilde{I}_1 - \left( \frac{\alpha_1}{K_1} \right) t \right)\\
+\frac{d\widetilde{I}_2}{dt} & = \left( \left( r_{22} K_2 \right) \widetilde{I}_2 
+			+ \left( r_{21} K_1 \right)  \widetilde{I}_1 
+			+ \left( r_{23} K_3 \right) \widetilde{I}_3 \right)
+		\left( 1 - \widetilde{I}_2 - \left( \frac{\alpha_2}{K_2} \right) t \right)\\
+\frac{d\widetilde{I}_3}{dt} & = \left( \left( r_{33} K_3 \right) \widetilde{I}_3 
+			+ \left( r_{32} K_2 \right)  \widetilde{I}_2 \right)
+		\left( 1 - \widetilde{I}_3 - \left( \frac{\alpha_3}{K_3} \right) t \right)
+\end{align*}
+
+If the second population is relatively small, then we might be able to 
+	decrease the total number of infections by focusing our entire efforts
+	on the smallest population to cut the link between the two populations.
+We then treat the vaccination problem as a pareto optimization problem, to distribute
+	the vaccines amongst the two populations.
+However this may not be better than splitting the vaccinations between the two.
+
 \section{Stochastic Model}
@@
 	Retrieved September 26, 2015
-	\url{http://www.wolframalpha.com/input/?i=ds%2Fdt+%3D+r+s+%28K+-+s+-+a+t%29}
-	\url{http://www.wolframalpha.com/input/?i=%E2%88%AB+e^%28-a+x^2+%2B+b+x%29+dx}
+	\url{http://www.wolframalpha.com/input/?i=\%E2\%88\%AB+e\^\%28-a+x\^2+\%2B+b+x\%29+dx}
+	\url{http://www.wolframalpha.com/input/?i=ds\%2Fdt+\%3D+r+s+\%28K+-+s+-+a+t\%29}
 \end{thebibliography}

### 2015-10-01 final_submission/manuscript.tex
updated sections

@@
 
-\section{Deterministic Model}
+\section{Deterministic SIR Model}
 \subsection{Algorithm and Assumptions}
@@
 The final parameter necessary to model vaccination strategies on a homogeneous population is the vaccination rate itself.
-When this is added, the system becomes:
+When this is added, the system becomes\cite{asu}:
 	\[
@@
 
+\section{Deterministically Allocating Vaccines Between Disjoint Populations}
+	Chris's stuff goes here
+
 \section{Stochastic Model}
-\subsection{Algorithm and Assumptions}
-	//TODO
-\subsection{Parameter Justifications}
-	//TODO
-\subsection{Predictions}
-	//TODO
-\subsection{Validation and Robustness Testing}
-	//TODO
-\subsection{Strengths and Weaknesses}
-	//TODO
-\subsection{Future Work}
-	//TODO
+	Paul's stuff goes here
 
@@
 		\url{http://web.stanford.edu/~jhj1/teachingdocs/Jones-on-R0.pdf}
+	\bibitem{asu}
+		\url{http://www.public.asu.edu/~hnesse/classes/sirv.html?Alpha=0.3&Beta=0.9&Nu=0.}
 	\end{thebibliography}

### 2015-10-01 proj1.tex
CONFLICTSS

--- 2015-09-30 proj1.tex
@@
 \usepackage{float}
-\usepackage{listings, color, times, textcomp, float, hyperref, subfigure}
+\usepackage{listings, color, times, textcomp, float, hyperref, subcaption}
 \definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
@@
 \includegraphics[width=\textwidth]{figures/Naive-model.png}
+\caption{Initial model predictions}
 \end{figure}
 The parts of each bars represent the respective number of citizens in a given age group that are sick. The black bars on the bottom represent how many people have been hospitalized so far.
-\par Unfortunately, research and according to a source from the CIC \cite{CIC} exposed many flaws in this model.
+\par Unfortunately, research and according to sources from the CIC \cite{CIC-stats} \cite{CIC-qa} exposed many flaws in this model.
 Mostly the fact that the odds of a given citizen staying sick for more than a month are extremely low, unlike in this model where they are actually quite high.
@@
 This also means that we can cut out an entire segment of the population, and generalize citizens younger than 25 years of age to a single slice.
-\par We also have more concrete data on how long citizens stay sick, and the likelihood of complications.
-Given that all data is given in terms of weeks, we change the model to compute the spread on a weekly basis.
-For simplicity and rounding, we assume that there are 4 weeks in a month, and that all citizens, regardless of age, have a 90\% of healing every week.
+As a slight additional adaptation, we also make it such that any given citizen is more likely to talk to people from their age slice.
+\par Secondly, we obtained more concrete data on how long citizens stay sick, and the likelihood of complications.
+The data we found was given in terms of weeks, so we changed the model to compute the spread on a weekly basis.
+For averaging and rounding, we assumed that there are 4 weeks in a month, and that all citizens, regardless of age, have a 90\% of healing every week.
 According to the CIC, a patient infected by the H1N1 virus is only contagious when showing symptoms.
 Therefore a citizen only infects other citizens when they are sick.
-Given what we know about the H1N1 virus and how contagious it is, we take 80\% to be the chance that a sick citizen infects a healthy one, with the 20\% accounting for sanitary precautions like hand sanitizer.
-\par
+Given what we know about the H1N1 virus and how contagious it is, we take 80\% to be the chance that a sick citizen infects a healthy one, with the 20\% accounting for sanitary precautions like hand sanitizer (assuming people don't have ``Swine flu Parties'' \cite{CIC-qa}).
+\par With this in mind, we adapted the model such that each age group had the attributes listed in Figure~\ref{fig:justified-values}:
 
@@
 	\begin{tabular}{c|c|c|c|c}
-Age group & Complication chance & Connectivity & Population percentage\\ \hline
-Junior & 0.05 & 100 & 100\%\\
-Adults & 0.1 & 50 & 50\%\\
-Seniors & 0.75 & 25 & 10\%
+		Age group & Heal rate & Complication chance & Connectivity & Population percentage\\ \hline
+		Junior & 0.9 & 0.05 & 25 & 60\%\\
+		Adults & 0.9 & 0.2 & 15 & 30\%\\
+		Seniors & 0.9 & 0.75 & 5 & 10\%
 	\end{tabular}
+	\caption{More justified attributes for age groups}
+	\label{fig:justified-values}
 \end{figure}
+
+Here are a handful of runs for different vaccination strategies:
+\begin{figure}
+	\centering
+	\begin{subfigure}[b]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{figures/Weekly-novax.png}
+		\caption{No vaccination policy}
+	\end{subfigure}~
+	\begin{subfigure}[b]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{figures/Weekly-randvax.png}
+		\caption{Random vaccination policy}
+	\end{subfigure}\\
+	\begin{subfigure}[b]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{figures/Weekly-juniors.png}
+		\caption{Prioritization of kids and teens}
+	\end{subfigure}~
+	\begin{subfigure}[b]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{figures/Weekly-seniors.png}
+		\caption{Prioritization of seniors}
+	\end{subfigure}
+	\caption{Model runs}
+	\label{fig:modelout}
+\end{figure}
+
+Which unfortunately doesn't look promising.
+
+\subsection{An explanation}
+As we saw in Figure~\ref{fig:modelout}
 
@@
 	\url{http://ithaca.areaconnect.com/statistics.htm}
-\bibitem{CIC}
+\bibitem{CIC-stats}
 	\url{http://www.cdc.gov/h1n1flu/surveillanceqa.htm}
+\bibitem{CIC-qa}
+	\url{http://www.cdc.gov/h1n1flu/qa.htm}
 \bibitem{walph}

### 2015-10-01 final_submission/manuscript.tex

@@
 \documentclass[titlepage]{article}
+\usepackage[margin=2.5cm]{geometry}
+\usepackage{enumerate}
+\usepackage{fancyhdr}
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{tikz}
+\usepackage{csvsimple}
+\usepackage{pgfplots}
 
-\usepackage{amsmath}
+\pagestyle{fancy}
+\fancyhead{}
+\fancyfoot{}
+\rfoot{\thepage}
+
+\usepackage{float}
+\usepackage{listings, color, times, textcomp, float, hyperref, subfigure}
+\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
+\definecolor{mylilas}{RGB}{170,55,241}
+\lstset{language=Matlab, basicstyle=\scriptsize\ttfamily,breaklines=true,frame=single,morekeywords={matlab2tikz},
+keywordstyle=\color{blue}, morekeywords=[2]{1}, keywordstyle=[2]{\color{black}}, identifierstyle=\color{black},
+stringstyle=\color{mylilas}, commentstyle=\color{mygreen}, showstringspaces=false, numbers=left,
+numberstyle={\tiny \color{black}}, numbersep=9pt, emph=[1]{for,end,break},emphstyle=[1]\color{red},
+literate={~} {\texttildelow}{1}}
+
+\newcommand{\exo}[1]{\section*{Exercise #1}}
+\newcommand{\prob}[1]{\section*{Problem #1}}
+\newcommand{\quest}[1]{\section*{Question #1}}
+\newcommand{\e}{&=}
+\newcommand{\p}[1]{\times 10^{#1}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrulewidth}{0.5pt}
+
+\DeclareMathOperator\erf{erf}
 
@@
 \date{September 30, 2015}
+
 \begin{document}
 \maketitle
+\thispagestyle{fancy}
 
@@
 \subsection{Algorithm and Assumptions}
-	Our initial research on epidemiology discovered a common model known as the SIR approach \cite{stanford}.
+	Our initial research on epidemiology discovered a common model known as the SIR approach \cite{SIR}.
 In this model, the population is segmented into three compartments that represent susceptible, infected, and removed states.
@@
 Now normalize these variables  to consider the fraction of the total population in each compartment, such that $s = S/N$, $i = I/N$, $r = R/N$.
-Then the model is governed by the following three ordinary differential equations \cite{stanford}:
+Then the model is governed by the following three ordinary differential equations \cite{SIR}:
 	\[
@@
 	\]
-	Here $\beta$ represents the effective contact rate, and $\gamma$ is the recovery rate \cite{stanford}.
+	Here $\beta$ represents the effective contact rate, and $\gamma$ is the recovery rate \cite{SIR}.
 It is important to note that in this variant of the SIR model, we interpreted R to stand for Recovered rather than removed.
@@
 This encompasses both individuals who recovered and those who died.
-We chose to seperate mortality rates out of the removal rate so that this parameter could be adjusted for the deadlier virus strain.
+We chose to separate mortality rates out of the removal rate so that this parameter could be adjusted for the deadlier virus strain.
 Then the system of equations becomes:
@@
 The final parameter necessary to model vaccination strategies on a homogeneous population is the vaccination rate itself.
-When this is added, the system becomes\cite{asu}:
+When this is added, the system becomes\cite{applet}:
 	\[
@@
 This does not depend on the fractions of susceptible or infected individuals because we have a constant $4000$ vaccines per month.
-In the actual implementation of this model, it was necessary to code in a conditional statement that set this constant to $0$ once the susceptible population completely diminished, otherewise this would result in negative population values for that compartment.
+In the actual implementation of this model, it was necessary to code in a conditional statement that set this constant to $0$ once the susceptible population completely diminished, otherwise this would result in negative population values for that compartment.
 Also, note that this assumes that the number of vaccines distributed grows linearly throughout the month, finally reaching $4000$ at the very end.\\
 
-	The final missing peice of this model was to do away with the well-mixed assumption.
+	The final missing piece of this model was to do away with the well-mixed assumption.
 This was crucial for us to make the distinction between highly-connected and frail individuals as requested by the government officials.
@@
 $\beta_{i,j}$ represents the rate of contact of an individual of type $i$ from an individual of type $j$.
-The values of these parameters come from a matrix called the ?
-that will be discussed in the ``Parameter Justifications'' section.
+The values of these parameters come from a matrix called the Next-Generation Matrix \cite{NGM} that will be discussed in the ``Parameter Justifications'' section.
 Finally, note that there is no subscript on the $\gamma$ parameter, meaning we assume the recovery rate to be constant for all four age group.
@@
 
-	We used MATLAB's built in ode45 solver on this system of equations to compute the fractions of suscpetible, infected, recovered, and dead individuals for each sub-segment of the population over time.
+	We used MATLAB's built in ode45 solver on this system of equations to compute the fractions of susceptible, infected, recovered, and dead individuals for each sub-segment of the population over time.
 \subsection{Parameter Justifications}
@@
 		\url{http://www.wolframalpha.com/input/?i=%E2%88%AB+e^%28-a+x^2+%2B+b+x%29+dx}
-	\bibitem{stanford}
-		\url{http://web.stanford.edu/~jhj1/teachingdocs/Jones-on-R0.pdf}
-	\bibitem{asu}
-		\url{http://www.public.asu.edu/~hnesse/classes/sirv.html?Alpha=0.3&Beta=0.9&Nu=0.}
+	\bibitem{SIR}
+		\url{http://web.SIR.edu/~jhj1/teachingdocs/Jones-on-R0.pdf}
+	\bibitem{applet}
+		\url{http://www.public.asu.edu/~hnesse/classes/sirv.html?}
+	\bibitem{NGM}
+		\url{http://rsif.royalsocietypublishing.org/content/7/47/873}
 	\end{thebibliography}

### 2015-10-01 proj1.tex
still many untracked files, it shouldn't compile

--- 2015-09-30 proj1.tex
@@
 
+\subsection{Improved Model}
+
+In the preceding model, once people became infected, they stayed
+	infected for the duration of the run.
+The following model adds a parameter, the ``curing rate'' $\gamma$,
+	which represents the rate at which individuals in the sick population
+	recover (and join the removed population).
+This model is the SIR model, and has standard parameters:
+
+\begin{align*}
+\frac{dI}{dt} & = \frac{ \beta S I}{N} - \gamma I\\
+\frac{dS}{dt} & = -\frac{\beta S I}{N} - r_v \\
+\frac{dR}{dt} & = r_v + \gamma I 
+\end{align*}
+
+We can use this model for pareto optimization.
+Cornell's population is about 21,000 \cite{cornellpop},
+	while ithaca's population is about 30,000 \cite{census}.
+Given a rate of vaccine distribution, by integrating the equations
+	we can determine how many total infections there are.
+If we had 4000 vaccinations to distribute, we can make a pareto
+	curve by plotting the set of points corresponding to different
+	relative distributions of the vaccinations.
+
+If one of the populations vaccinates all of its seceptible individuals
+	before the other, we need to decide what to do with the vaccines
+	that it is not consuming.
+If we choose not to redistribute them, the pareto curve looks like this:
+
+\includegraphics[width=0.5\textwidth]{figures/sir-pareto.pdf}
+
+If we choose to redistribute them, for $\beta = 6$ and $\gamma = 4$,
+	the curve looks like this:
+
+\includegraphics[width=0.5\textwidth]{figures/sir-pareto-top.pdf}
+
+And the pareto curve looks like:
+
+\includegraphics[width=0.5\textwidth]{figures/sir-pareto-top-pareto.pdf}
+
+We conclude that making sure that all of the vaccines are allocated is very
+	important for any vaccine model, noting the difference.
+The pareto-optimal points are when 71-84\% of the vaccine goes to the
+	cornellians and 29-16\% goes to the ithacans, 
+	and when 15-20\% of the vaccine goes to the cornellians 
+	and 80-85\% of the vaccine goes to the ithacans.
+
 \subsection{Disease Spread between Coupled Populations}
@@
 	\url{http://www.wolframalpha.com/input/?i=ds\%2Fdt+\%3D+r+s+\%28K+-+s+-+a+t\%29}
+\bibitem{census}
+	http://quickfacts.census.gov/qfd/states/36/3638077.html
+\bibitem{cornellpop}
+	https://www.cornell.edu/about/facts.cfm
 \end{thebibliography}

### 2015-10-01 models/deterministic/evaluate-two-way-splitting.m
mega-commit!!!!!!

@@ -1,0 +1,12 @@
+global n_staff = 6000;
+global n_students = 25000;
+global n_ithacans = 22000;
+
+% entry 
+global beta = ;
+global gamma = 
+
+function derivs = f(t, x)
+	global n_staff; global n_students; global n_ithacans;
+	I = x(1:3); S = x(4:6); R = x(7:9);
+	

### 2015-10-01 models/deterministic/numerical_sir/mod3.m

@@ -1,0 +1,69 @@
+clc; close all; clear;
+%=======================================================%
+% SIR model with death, vaccines, and 4 sub-populations %
+%=======================================================%
+pkg load odepkg;
+
+% ----- Model Parameters ----- %
+N = 60000;
+N_students = 21000;
+N_staff = 9000;
+N_residents = N - N_students - N_staff;
+demographics = [ N_students; N_staff; N_residents] / N;
+s0 = .95 * demographics;
+i0 = .05 * demographics;
+r0 = .0 * demographics;
+d0 = .0 * demographics;
+
+contagious_time = .25;
+R0 = [3.00, 2.00  0.10; ...
+      2.00, 2.00, 1.00; ...
+      0.10, 1.00, 1.50];
+beta = R0 / contagious_time;
+mu = 0.00001709 * [4.5;2.1;1.0;1.7]; % * 20;
+vaccines = 4000;
+vaccine_distribution = [0.3; 0.3; 0.4];
+gamma = vaccines*vaccine_distribution/N;
+% ---------------------------- %
+
+t0 = 0;
+tf = 8;
+y0 = [s0;i0;r0;d0];
+[T,Y] = ode45(@(t,y) sir_update_c(y, beta, nu, mu, gamma), [t0, tf], y0)
+
+
+
+%	titles = {'Students','Staff','Ithacans'};
+%	for i = 1:4
+%	    figure
+%	    switch i
+%	        case 1
+%	            set(gcf,'units','normalized','outerposition',[0 0.5 .5 .5])
+%	        case 2
+%	            set(gcf,'units','normalized','outerposition',[0.5 0.5 .5 .5])
+%	        case 3
+%	            set(gcf,'units','normalized','outerposition',[0 0 .5 .5])
+%	        otherwise
+%	            set(gcf,'units','normalized','outerposition',[0.5 0 .5 .5])
+%	    end
+%	    plot(T,Y(:,i:4:8+i) / demographics(i))
+%	    ylim([0,1])
+%	    title(titles(i))
+%	    xlabel('Months since outbreak')
+%	    ylabel('Fraction of population')
+%	    legend('Susceptible', 'Infected', 'Removed')
+%	end
+%	
+%	total_deaths = zeros(1,4);
+%	epidemic_duration = zeros(1,4);
+%	total_suffering = zeros(1,4);
+%	
+%	for i = 1:4
[…]

### 2015-10-01 proj1.tex
Adding justifications of gamma

@@
 	recover (and join the removed population).
+This will be the inverse of the time spent sick, which is on average 
+	1 week \cite{CIC-stats}.
+The value of the interaction parameter $\beta$ is determined in \textbf{PLACEHOLDER}
 This model is the SIR model, and has standard parameters:
@@
 
-\subsection{Disease Spread between Coupled Populations}
+\subsection{Coupling the Populations}
 
-Consider three populations, with two connected only by an intermediate
-	one.
+Consider the following modifications to the model, to add interactions.
+Let $\alpha$ be the interaction coefficient.
+$\alpha$ should be approximately a fraction of $\beta$.
+For example, if a member of one population is one tenth as likely
+	to interact with the other as with one from their own population,
+	the value for $\alpha$ would be $\frac{\beta}{10}$.
+It represents how likely the seceptible of $i$ are to be infected by the
+	infected of $j$.
 
 \begin{align*}
-\frac{d\widetilde{I}_1}{dt} & = \left( \left( r_{11} K_1 \right) \widetilde{I}_1 
-			+ \left( r_{12} K_2 \right)  \widetilde{I}_2 \right)
-		\left( 1 - \widetilde{I}_1 - \left( \frac{\alpha_1}{K_1} \right) t \right)\\
-\frac{d\widetilde{I}_2}{dt} & = \left( \left( r_{22} K_2 \right) \widetilde{I}_2 
-			+ \left( r_{21} K_1 \right)  \widetilde{I}_1 
-			+ \left( r_{23} K_3 \right) \widetilde{I}_3 \right)
-		\left( 1 - \widetilde{I}_2 - \left( \frac{\alpha_2}{K_2} \right) t \right)\\
-\frac{d\widetilde{I}_3}{dt} & = \left( \left( r_{33} K_3 \right) \widetilde{I}_3 
-			+ \left( r_{32} K_2 \right)  \widetilde{I}_2 \right)
-		\left( 1 - \widetilde{I}_3 - \left( \frac{\alpha_3}{K_3} \right) t \right)
+\frac{dI_1}{dt} & = \frac{ \beta S_1 I_1}{N_1} + \frac{ \alpha S_1 I_2 }{N_2} - \gamma I_1\\
+\frac{dS_1}{dt} & = -\frac{\beta S_1 I_1}{N_1}  - \frac{ \alpha S_1 I_2 }{N_2} - r_v \\
+\frac{dR_1}{dt} & = r_v + \gamma I_1 \\
+\frac{dI_2}{dt} & = \frac{ \beta S_2 I_2}{N}  + \frac{ \alpha S_2 I_1 }{N_1} - \gamma I_2\\
+\frac{dS_2}{dt} & = -\frac{\beta S_2 I_2}{N}  - \frac{ \alpha S_2 I_1 }{N_1} - r_v \\
+\frac{dR_2}{dt} & = r_v + \gamma I_2 
 \end{align*}
 
-If the second population is relatively small, then we might be able to 
-	decrease the total number of infections by focusing our entire efforts
-	on the smallest population to cut the link between the two populations.
-We then treat the vaccination problem as a pareto optimization problem, to distribute
-	the vaccines amongst the two populations.
-However this may not be better than splitting the vaccinations between the two.
+This model, in the fully interacting case, produces the following pareto curve:
+
+\includegraphics[width=0.5\textwidth]{figures/sir-interacting.pdf}
+
+This doesn't show a significant variance across different immunization strategies.
+It seems like vaccinating one population, and then the other, is a reasonable
+	immunization strategy.
+If we want to be in the middle, that is alright.
 

### 2015-10-01 proj1.tex
Final commit

--- 2015-10-01 proj1.tex
@@
 Then, to model the spread, we associated arbitrary recovery rates, complication rates and number of people talked to per month to each age group, giving us the following table:
-\begin{figure}[H]
+\begin{figure}[h!]
 \centering
@@
 \end{tabular}
+\caption{Initial citizen attributes}
 \end{figure}
@@
 
-\begin{figure}[H]
+\begin{figure}[h!]
 	\centering
@@
 Here are a handful of runs for different vaccination strategies:
-\begin{figure}
+\begin{figure}[h!]
 	\centering
@@
 
-\subsection{An explanation}
-As we saw in Figure~\ref{fig:modelout}
+\subsection{Analysis}
+\par As we saw in Figure~\ref{fig:modelout}, vaccination policies don't actually impact the spread of the disease too much.
+Rather, it seems to be more helpful for damage control.
+We see from the 4 graphs in Figure~\ref{fig:modelout} the peak infection rate seems to stay near 10\% regardless of the vaccination policy.
+On the other hand, it seems to help for damage control because even though the infection rate stays the same, when we vaccinate more seniors, fewer people end up hospitalized than when we target more juniors.
+But surprisingly, the random vaccination policy consistently came out on top.
+This peaked our interest, so we ran many simulations comparing the three models.
+Figure~\ref{fig:10outs} shows the outputs side by side.
+We see that the random policy actually puts just about as many people in the hospital as the policy that targets seniors, while keeping the peak infection rate actually lower than all the policies.
+
+\begin{figure}[h!]
+\centering
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\begin{tabular}{rc|c}
+& Peak infection (\%) & Total hospitalizations\\ \cline{2-3}
+& 11.1  & 431\\
+& 10.62 & 418\\
+& 9.47  & 433\\
+& 8.9   & 414\\
+& 11.11 & 404\\
+& 10.24 & 411\\
+& 11.35 & 442\\
+& 10.59 & 412\\
+& 10.88 & 432\\
+& 10    & 395\\ \hline
+Mean & 10.426 & 419.2
+\end{tabular}
+\caption{Juniors first}
+\end{subfigure}~
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\begin{tabular}{rc|c}
+& Peak infection (\%) & Total hospitalizations\\ \cline{2-3}
+& 10.45 & 389\\
+& 10.65 & 376\\
+& 10.25 & 419\\
+& 10.86 & 388\\
+& 10.1  & 384\\
+& 11.05 & 373\\
+& 9.97  & 377\\
+& 9.57  & 383\\
+& 10.14 & 363\\
+& 9.91  & 367\\ \hline
+Mean & 10.295 & 381.9
+\end{tabular}
+\caption{Seniors first}
+\end{subfigure}\\
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\begin{tabular}{rc|c}
+& Peak infection (\%) & Total hospitalizations\\ \cline{2-3}
+& 9.5   & 378\\
+& 9.99  & 393\\
+& 10.02 & 373\\
+& 9.28  & 370\\
+& 9.72  & 386\\
+& 9.68  & 369\\
+& 10.25 & 423\\
+& 10.36 & 415\\
+& 9.11  & 370\\
+& 9.96  & 352\\ \hline
+Mean & 9.787 & 382.9
+\end{tabular}
+\caption{Indiscriminate vaccinations}
+\end{subfigure}
+\caption{Comparison of 10 runs for different vaccination policies}
+\label{fig:10outs}
+\end{figure}
 

### 2015-10-01 models/stochastic/spread.m

@@ -89,22 +89,22 @@ function [peak, hosps, sick_per_delta, hospitalized] = spread(age_range, p)
   for w = 1:weeks
     disp(sprintf('Week: %d', w));
     vax = 0;
-    for k = ifelse(age_range == 1, 1:ithaca_pop, ithaca_pop:-1:1)
-      if ~vaccinated(k)
-        vaccinated(k) = true;
-        vax = vax + 1;
-      end
-      if vax == vaccines
-        break
-      end
-    end
-    % while vax ~= vaccines
-    %   r = round(rand*(ithaca_pop-1))+1;
-    %   if ~vaccinated(r)
-    %     vaccinated(r) = true;
+    % for k = ifelse(age_range == 1, 1:ithaca_pop, ithaca_pop:-1:1)
+    %   if ~vaccinated(k)
+    %     vaccinated(k) = true;
     %     vax = vax + 1;
     %   end
+    %   if vax == vaccines
+    %     break
+    %   end
     % end
+    while vax ~= vaccines
+      r = round(rand*(ithaca_pop-1))+1;
+      if ~vaccinated(r)
+        vaccinated(r) = true;
+        vax = vax + 1;
+      end
+    end
     for c = 1:ithaca_pop
       if sick(c)
         if (rand < heal_chance)

### 2015-10-01 final_submission/manuscript.tex
any changes?

@@
 \subsection{Parameter Justifications}
-	//TODO
+	An important dependent parameter for modeling epidemics is known as the Basic Reproduction rate \cite{SIR}. This quantity is:
+	\[
+		R_0 = r c d
+	\]
+	Where $r$ is the probability of an infection given contact between a susceptible and infected individual, $c$ is the average rate of contact between susceptible and infected individuals, and $d$ is the duration of infectiousness \cite{SIR}.
+This $R_0$ value can be used to compute our $\beta$ value in the SIR model.
+Because we need the number of infected individuals to increase for the basic SIR model, we have:
+	\[
+		\frac{di}{dt} = \beta s i - \gamma i > 0
+	\]
+	\[
+		\frac{\beta s i}{\gamma} > i
+	\]
+	\begin{center}
+		$s \approx 1$ at the start of the epidemic
+	\end{center}
+	\[
+		\frac{\beta}{\gamma} = R_0 > 1
+	\]
+	This is then the relationship between $\beta$ and $R_0$ \cite{SIR}.
+We can compute $\gamma$ if we define it to be $1/d$, the duration of infectiousness \cite{SIR}.
+The fact that we assume this duration of infectiousness to be the same across all ages is why we see a constant recovery rate $\gamma$.The average duration of infectiousness for swine flu is one week \cite{oneweek}(.25 months), and the Basic Reproduction Number is 1.5 \cite{R0}, so we have $\gamma = 4$ and $\beta = 6$ in the standard SIR model.
+When data was first being collected in the 2009 outbreak of the swine flu, approximately 7\% of the population was infected \cite{init}, so we began our SIR model at s = .95 and r = .05.
+Finally, in order to set the death rate for the homogeneous SIR model, we discovered that of approximately 60.8 million infections in 2009, 12,469 resulted in deaths \cite{deathpercent}.
+Dividing this fraction between 12 months leads to $\mu = 0.00001709$
+
+    For the more complicated SIR model, however, single parameter values do not work because we have four distinct sub-populations. These sub-populations were chosen based on the following bar chart:
+
+	\begin{figure}[h!]
+	\centering
+		\includegraphics[width=.5\textwidth]{figures/death_graph.png}
+        \caption{Mortality rates by age group \cite{deathgraph}}
+	\end{figure}
+
+    Here you can see that the death rate is relatively constant from age 25-64, so we grouped these ranges together into the ``adults'' category.
+The $\mu$ value for each sub-population was simply these numbers scaled by the global $\mu$ found above.
+The fractions of the total population that each of these sub-groups occupied \cite{pop} was found to be (.0248, .6058, 3066, .0628).
+Finally, the single $\beta$ had to be derived from a Next Generation Matrix.
+This is an equivalent of the $B0$ value for multiple sub-populations with transitions between each.
+This is defined to be that the jth column represents the average number of infections of individuals of type i from an individual of type j \cite{SIR}.
+These values were largely hand-picked for the 4 sub-population model, but we tried not to deviate too far from 1.5.
+    \[
+        R0 = 
+        \begin{bmatrix}
+        3.00 & 0.10 & 0.10 & 0.10 \\
+        0.10 & 3.00 & 1.00 & 0.10 \\
+        2.00 & 2.00 & 1.50 & 1.00 \\
+        0.10 & 0.10 & 0.50 & 1.00
+        \end{bmatrix}
+    \]
+
 \subsection{Predictions}
 	//TODO
-\subsection{Validation and Robustness Testing}
-	//TODO
-\subsection{Strengths and Weaknesses}
-	//TODO
-\subsection{Future Work}
-	//TODO
+% \subsection{Validation and Robustness Testing}
+% 	//TODO
+% \subsection{Strengths and Weaknesses}
+% 	//TODO
+% \subsection{Future Work}
+% 	//TODO
 
@@
 
-\section{Conclusions}
+\section{Conclusion}
 	//TODO
@@
 	\bibitem{SIR}
-		\url{http://web.SIR.edu/~jhj1/teachingdocs/Jones-on-R0.pdf}
+		\url{http://web.stanford.edu/~jhj1/teachingdocs/Jones-on-R0.pdf}
 	\bibitem{applet}
@@
 		\url{http://rsif.royalsocietypublishing.org/content/7/47/873}
+	\bibitem{oneweek}
+		\url{https://www.cdph.ca.gov/HealthInfo/h1n1flufaqs/Pages/H1N1fluFAQs-01-GenInfo.aspx}
+	\bibitem{R0}
+		\url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2715422/}
+	\bibitem{deathgraph}
+		\url{http://www.cdc.gov/h1n1flu/surveillanceqa.htm}
+    \bibitem{init}
+        \url{http://academic.udayton.edu/muhammadusman/2010Stander/H1N1.pdf}
+    \bibitem{deathpercent}
+        \url{http://www.cdc.gov/h1n1flu/estimates_2009_h1n1.htm}
+    \bibitem{pop}
+        \url{http://ithaca.areaconnect.com/statistics.htm}
 	\end{thebibliography}

### 2015-10-01 final_submission/manuscript.tex
Mostly merged

@@
 \section{Deterministically Allocating Vaccines Between Disjoint Populations}
-	Chris's stuff goes here
+
+First, we propose a deterministic model which doesn't divide the population
+	into sections at all.
+This model will help give us an idea of what the effect of dividing
+	the population into sections will be.
+In addition, it can give us an idea of how vaccination affects the spread of
+	disease within a population.
+
+\subsection{Disease Spread with No Vaccine}
+
+We include this derivation of the logistic model for disease spread
+	to compare with our derivation of the modified logistic model
+	for disease spread and vaccination.
+
+Suppose there are two groups of people, with $I$ representing the number of
+	infected people, and $S$ represnting the number of people seceptible for the
+	disease.
+Suppose the probability of a single healthy person becoming sick is proportional
+	only to the total number of sick people.
+Since there are $S$ healthy people, the expected number of healthy people
+	who become sick should be proportional to $S I$.
+This model does not take into account recovery from sickness.
+If we introduce a parameter $r$, we can express this relationship as a system
+	of differential equations:
+
+\begin{align*}
+\frac{dS}{dt} & = - r S I \\
+\frac{dI}{dt} & = r S I \\
+\end{align*}
+
+Notice that the total number of people, $S + I$, is conserved, i.e.,
+
+\[ \frac{d}{dt} \left( S + I \right) = \frac{dS}{dt} + \frac{dI}{dt}
+	= r S I - r S I = 0 \]
+
+If we introduce another parameter, $K$, to represent the total population,
+	such that $S + I = K$ for all time, then the differential equations
+	can be decoupled:
+
+\begin{align*}
+\frac{dS}{dt} & = - r S (K - S) \\
+\frac{dI}{dt} & = r (K - I) I \\
+\end{align*}
+
+The two differential equations shown here are now two independant, uncoupled
+	differential equations which we can solve seperately.
+
+\subsubsection{Nondimensionalizing the Model}
+
+It would be helpful to make all of the constants dimensionless parameters.
+This will give us insight into the different scales that the model works at.
+
+To turn $I$ into a dimensionless parameter, we can divide by $K$.
+
+\begin{align*}
+\frac{d\left(\frac{I}{K}\right)}{dt}
+	& = \left( r K \right) \left( \frac{I}{K} \right) \left( 1 - \frac{I}{K} \right)\\
+\frac{d\widetilde{I}}{dt}
+	& = \widetilde{r} \widetilde{I} \left( 1 - \widetilde{I} \right)
+\end{align*}
+
+This shows that the dynamics of $I$ do not depend on its size.
+For each value of the dimensionless constant $\widetilde{r}$,
+	we determine the behavior from it.
+
+\subsection{Diease Spread with Vaccine}
+
+Suppose we introduce a third group, $V$ for vaccinate, representing people who are
+	immune from the diesease.
+Right now, the only way to become immune is through vaccination,
+	but we may add recovery in subsequent models.
+Suppose we randomly vaccinate members of the healthy population,
+	a constant number each time step.
+This introduces a third parameter, $r_v$, the rate of vaccinations.
+The vaccinated population isn't seceptible to the disease.
+This process is modeled by the following system of diffential equations:
+
+\begin{align*}
+\frac{dI}{dt} & = r S I \\
+\frac{dS}{dt} & = - r S I - r_v \\
+\frac{dV}{dt} & = r_v
+\end{align*}
+
+This model is accurate only if all three of the populations remain
+	positive.
+This is important to remember.
+The number of healthy people will always be decreasing, so once
+	the number of healthy people hits zero, we stop the model.
+The sick people stay sick, and the vaccinated people stay vaccinated.
+
+The number of vaccinated people as a function of time is:
+
+\[ V(t) = r_v t \]
+
+Again, we find the conserved quantity $\left( S + I + V \right)$ and
+	name it $K$.
+For the model to be valid, we need $S = K - I - V \geq 0$.
+Therefore, the model is valid as long as:
+
+\[ K \geq I + r_v t \]
+
+\includegraphics[width=0.5\textwidth]{figures/admissible-region.pdf}
+
+In the above figure, once $I(t)$ hits the diagonal line,
+	it stays horizontal and keeps its value.
+
+We have now made these differential equations independant,
+	within the model's zone of validity:
+
+\begin{align*}
+\frac{dI}{dt} & = r I \left( K - I - r_v t \right)\\
+\frac{dS}{dt} & = - r S \left( K - S - r_v t \right) - r_v \\
+\frac{dV}{dt} & = r_v t
+\end{align*}
+
+We present a solution for $I(t)$ \cite{walph}:
+
+\begin{align}
+I(t) & =
+	\frac{e^{ r t (K - \frac{ r_v t}{2})}}
+	{\frac{1}{S(0)} + \sqrt{\frac{\pi r}{2 r_v}}
+		e^{\frac12 \left( \frac{r}{r_v} \right) K^2}
+		\left(\erf\left(  \sqrt{\frac{r}{2 r_v}} (r_v t - K ) \right)
+			- \erf\left( - \sqrt{\frac{r}{2 r_v}} K \right) \right)}
+\end{align}
+
+From the above expression for the boundary of the valid region,
+	we get the following implicit expression for $t_f, I(t_f)$
+
+\begin{align}
+I(t_f) =
+	\frac{e^{ r t_f (K - \frac{ r_v t_f }{2})}}
+	{\frac{1}{S(0)} + \sqrt{\frac{\pi r}{2 r_v}}
+		e^{\frac12 \left( \frac{r}{r_v} \right) K^2}
+		\left(\erf\left(  \sqrt{\frac{r}{2 r_v}} (r_v t_f - K ) \right)
+			- \erf\left( - \sqrt{\frac{r}{2 r_v}} K \right) \right)}
+	& = K - r_v t_f
+\end{align}
+
+Here is a family of values of $I(t)$.
+The total population and the rate of distributing vaccines are
+	held constant, while each curve corresponds to a different
+	value of the growth rate constant.
+Each curve starts with an initial condition at 5 percent of the
+	population, and the growth rate determines how many people
+	overall are infected.
+
+\includegraphics[width=0.5\textwidth]{figures/vaccination-model-curves-varying-s.pdf}
+
+\subsection{Pareto Optimizing Between Two Independent Populations}
+
+The model above modeled how one population responds to a combination
+	of disease spreading and vaccines.
+The model depends on $K$, the total population, $I(0)$, the initial
+	number of sick people, $r$, the rate of infection growth,
+	and $r_v$, the vaccination rate.
+Of these parameters, in a real world setting, we can expect to know $K$
+	very precicely from census data, approximate $I(0)$ as small,
+	have very little idea about $r$, and control $r_v$.
+
+I want to extend this problem to two independant populations.
+Each population will get different sets of constants,
+	$K_1$ and $K_2$, $I_1(0)$ and $I_2(0)$, $r_1$ and $r_2$.
+The goal will be to figure out how to optimally allocate the
+	vaccines between the two populations.
+The decision variable, $d$, will be the fraction of the total vaccination
+	rate I allocate to each population.
+Population 1 will be vaccinated at a rate $d r_v$, while population
+	2 will be vaccinated at a rate $(1 - d) r_v$.
+In this model, I try to minimize the final number of sick people in
+	each population.
+This will give me a pareto front.
+
+The following collection of pareto fronts comes from keeping the disease
+	growth rate the same for both, and varying its magnitude.
+
+\includegraphics[width=0.5\textwidth]{figures/vaccination-model-same-r-pareto-curves.pdf}
+
+This collection of pareto fronts comes from keeping the growth
+	rate for population 2 the same, and varying the growth
+	rate of population 1.
+
+\includegraphics[width=0.5\textwidth]{figures/vaccination-model-different-r-pareto-curves.pdf}
+
+\subsection{Improved Model}
+
+In the preceding model, once people became infected, they stayed
+	infected for the duration of the run.
+The following model adds a parameter, the ``curing rate'' $\gamma$,
+	which represents the rate at which individuals in the sick population
+	recover (and join the removed population).
+This will be the inverse of the time spent sick, which is on average
+	1 week \cite{CIC-stats}.
+The value of the interaction parameter $\beta$ is determined in \textbf{PLACEHOLDER}
+This model is the SIR model, and has standard parameters:
+
+\begin{align*}
+\frac{dI}{dt} & = \frac{ \beta S I}{N} - \gamma I\\
+\frac{dS}{dt} & = -\frac{\beta S I}{N} - r_v \\
+\frac{dR}{dt} & = r_v + \gamma I
+\end{align*}
+
+We can use this model for pareto optimization.
+Cornell's population is about 21,000 \cite{cornellpop},
+	while ithaca's population is about 30,000 \cite{census}.
+Given a rate of vaccine distribution, by integrating the equations
+	we can determine how many total infections there are.
+If we had 4000 vaccinations to distribute, we can make a pareto
+	curve by plotting the set of points corresponding to different
+	relative distributions of the vaccinations.
+
+If one of the populations vaccinates all of its seceptible individuals
+	before the other, we need to decide what to do with the vaccines
+	that it is not consuming.
+If we choose not to redistribute them, the pareto curve looks like this:
+
+\includegraphics[width=0.5\textwidth]{figures/sir-pareto.pdf}
+
+If we choose to redistribute them, for $\beta = 6$ and $\gamma = 4$,
+	the curve looks like this:
+
+\includegraphics[width=0.5\textwidth]{figures/sir-pareto-top.pdf}
+
+And the pareto curve looks like:
+
+\includegraphics[width=0.5\textwidth]{figures/sir-pareto-top-pareto.pdf}
+
+We conclude that making sure that all of the vaccines are allocated is very
+	important for any vaccine model, noting the difference.
+The pareto-optimal points are when 71-84\% of the vaccine goes to the
+	cornellians and 29-16\% goes to the ithacans,
+	and when 15-20\% of the vaccine goes to the cornellians
+	and 80-85\% of the vaccine goes to the ithacans.
+
+\subsection{Coupling the Populations}
+
+Consider the following modifications to the model, to add interactions.
+Let $\alpha$ be the interaction coefficient.
+$\alpha$ should be approximately a fraction of $\beta$.
+For example, if a member of one population is one tenth as likely
+	to interact with the other as with one from their own population,
+	the value for $\alpha$ would be $\frac{\beta}{10}$.
+It represents how likely the seceptible of $i$ are to be infected by the
+	infected of $j$.
+
+\begin{align*}
+\frac{dI_1}{dt} & = \frac{ \beta S_1 I_1}{N_1} + \frac{ \alpha S_1 I_2 }{N_2} - \gamma I_1\\
+\frac{dS_1}{dt} & = -\frac{\beta S_1 I_1}{N_1}  - \frac{ \alpha S_1 I_2 }{N_2} - r_v \\
+\frac{dR_1}{dt} & = r_v + \gamma I_1 \\
+\frac{dI_2}{dt} & = \frac{ \beta S_2 I_2}{N}  + \frac{ \alpha S_2 I_1 }{N_1} - \gamma I_2\\
+\frac{dS_2}{dt} & = -\frac{\beta S_2 I_2}{N}  - \frac{ \alpha S_2 I_1 }{N_1} - r_v \\
+\frac{dR_2}{dt} & = r_v + \gamma I_2
+\end{align*}
+
+This model, in the fully interacting case, produces the following pareto curve:
+
+\includegraphics[width=0.5\textwidth]{figures/sir-interacting.pdf}
+
+This doesn't show a significant variance across different immunization strategies.
+It seems like vaccinating one population, and then the other, is a reasonable
+	immunization strategy.
+If we want to be in the middle, that is alright.
 
 \section{Stochastic Model}
-	Paul's stuff goes here
+\subsection{A naive model}
+\par At first, we made a simple, arbitrary stochastic model just to get a rough idea of how the disease would spread.
+Over the course of 5 months, we would look at the spread of the disease given certain conditions.
+Every citizen fell into an age category, according to demographic figures from \emph{areaConnect} \cite{areconnect}.
+Then, to model the spread, we associated arbitrary recovery rates, complication rates and number of people talked to per month to each age group, giving us the following table:
+\begin{figure}[h!]
+\centering
+\begin{tabular}{c|c|c|c|c}
+Age group & Heal rate & Complication chance & Connectivity & Population percentage\\ \hline
+Kids & 0.5 & 0.2 & 20 & 10\%\\
+Teens & 0.4 & 0.1 & 30 & 40\%\\
+Adults & 0.3 & 0.3 & 25 & 40\%\\
+Seniors & 0.2 & 0.7 & 10 & 10\%
+\end{tabular}
+\caption{Initial citizen attributes}
+\end{figure}
+
+\par The Heal rate and Complication chance columns represent the odds of a given citizen to recover from the disease in a month.
+The Connectivity column represents the number of people each sick citizen has a chance to infect, uniformly sampled from the entire population.
+We would then infect 500 people initially, and let the simulation run its course.
+In this particular example, we vaccinate randomly.
+We also assume that a person who has been sick cannot be sick again, and that when complications arise in a given patient, they are taken to the hospital and do not keep infecting the rest of the population.
+Here's how the spread of disease looks:
+\begin{figure}[h!]
+\includegraphics[width=\textwidth]{figures/Naive-model.png}
+\caption{Initial model predictions}
+\end{figure}
+The parts of each bars represent the respective number of citizens in a given age group that are sick. The black bars on the bottom represent how many people have been hospitalized so far.
+\par Unfortunately, research and according to sources from the CIC \cite{CIC-stats} \cite{CIC-qa} exposed many flaws in this model.
+Mostly the fact that the odds of a given citizen staying sick for more than a month are extremely low, unlike in this model where they are actually quite high.
+In fact, every age range stays sick for just about the same amount of time, with seniors sick for 2 weeks.
+The CIC data also exposed something very interesting, which is that seniors are not the ones in the most danger from the flu, kids smaller than the age of 5 are.
+This gave us reason to use a model that looks at the spread of the disease on a weekly basis.
+
+\subsection{A more justified approach}
+\par As mentioned in earlier, the CIC data gave us a lot to think about in terms of refining this model.
+First of all, we looked at the effects of the population of children in Ithaca, where children are citizens younger than 5 years of age.
+Since their risks of complications are so much higher, a good idea would be to immediately vaccinate all of them to avoid a devastating death toll.
+Fortunately, according to \emph{areaConnect}, there are less than 2,000 children in Ithaca, meaning that they can all be vaccinated as soon as the shipment of vaccines arrives.
+This also means that we can cut out an entire segment of the population, and generalize citizens younger than 25 years of age to a single slice.
+As a slight additional adaptation, we also make it such that any given citizen is more likely to talk to people from their age slice.
+\par Secondly, we obtained more concrete data on how long citizens stay sick, and the likelihood of complications.
+The data we found was given in terms of weeks, so we changed the model to compute the spread on a weekly basis.
+For averaging and rounding, we assumed that there are 4 weeks in a month, and that all citizens, regardless of age, have a 90\% of healing every week.
+According to the CIC, a patient infected by the H1N1 virus is only contagious when showing symptoms.
+Therefore a citizen only infects other citizens when they are sick.
+Given what we know about the H1N1 virus and how contagious it is, we take 80\% to be the chance that a sick citizen infects a healthy one, with the 20\% accounting for sanitary precautions like hand sanitizer (assuming people don't have ``Swine flu Parties'' \cite{CIC-qa}).
+\par With this in mind, we adapted the model such that each age group had the attributes listed in Figure~\ref{fig:justified-values}:
+
+\begin{figure}[h!]
+	\centering
+	\begin{tabular}{c|c|c|c|c}
+		Age group & Heal rate & Complication chance & Connectivity & Population percentage\\ \hline
+		Junior & 0.9 & 0.05 & 25 & 60\%\\
+		Adults & 0.9 & 0.2 & 15 & 30\%\\
+		Seniors & 0.9 & 0.75 & 5 & 10\%
+	\end{tabular}
+	\caption{More justified attributes for age groups}
+	\label{fig:justified-values}
+\end{figure}
+
+Here are a handful of runs for different vaccination strategies:
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}[b]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{figures/Weekly-novax.png}
+		\caption{No vaccination policy}
+	\end{subfigure}~
+	\begin{subfigure}[b]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{figures/Weekly-randvax.png}
+		\caption{Random vaccination policy}
+	\end{subfigure}\\
+	\begin{subfigure}[b]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{figures/Weekly-juniors.png}
+		\caption{Prioritization of kids and teens}
+	\end{subfigure}~
+	\begin{subfigure}[b]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{figures/Weekly-seniors.png}
+		\caption{Prioritization of seniors}
+	\end{subfigure}
+	\caption{Model runs}
+	\label{fig:modelout}
+\end{figure}
+
+Which unfortunately doesn't look promising.
+
+\subsection{Analysis}
+\par As we saw in Figure~\ref{fig:modelout}, vaccination policies don't actually impact the spread of the disease too much.
+Rather, it seems to be more helpful for damage control.
+We see from the 4 graphs in Figure~\ref{fig:modelout} the peak infection rate seems to stay near 10\% regardless of the vaccination policy.
+On the other hand, it seems to help for damage control because even though the infection rate stays the same, when we vaccinate more seniors, fewer people end up hospitalized than when we target more juniors.
+But surprisingly, the random vaccination policy consistently came out on top.
+This peaked our interest, so we ran many simulations comparing the three models.
+Figure~\ref{fig:10outs} shows the outputs side by side.
+We see that the random policy actually puts just about as many people in the hospital as the policy that targets seniors, while keeping the peak infection rate actually lower than all the policies.
+
+\begin{figure}[h!]
+\centering
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\begin{tabular}{rc|c}
+& Peak infection (\%) & Total hospitalizations\\ \cline{2-3}
+& 11.1  & 431\\
+& 10.62 & 418\\
+& 9.47  & 433\\
+& 8.9   & 414\\
+& 11.11 & 404\\
+& 10.24 & 411\\
+& 11.35 & 442\\
+& 10.59 & 412\\
+& 10.88 & 432\\
+& 10    & 395\\ \hline
+Mean & 10.426 & 419.2
+\end{tabular}
+\caption{Juniors first}
+\end{subfigure}~
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\begin{tabular}{rc|c}
+& Peak infection (\%) & Total hospitalizations\\ \cline{2-3}
+& 10.45 & 389\\
+& 10.65 & 376\\
+& 10.25 & 419\\
+& 10.86 & 388\\
+& 10.1  & 384\\
+& 11.05 & 373\\
+& 9.97  & 377\\
+& 9.57  & 383\\
+& 10.14 & 363\\
+& 9.91  & 367\\ \hline
+Mean & 10.295 & 381.9
+\end{tabular}
+\caption{Seniors first}
+\end{subfigure}\\
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\begin{tabular}{rc|c}
+& Peak infection (\%) & Total hospitalizations\\ \cline{2-3}
+& 9.5   & 378\\
+& 9.99  & 393\\
+& 10.02 & 373\\
+& 9.28  & 370\\
+& 9.72  & 386\\
+& 9.68  & 369\\
+& 10.25 & 423\\
+& 10.36 & 415\\
+& 9.11  & 370\\
+& 9.96  & 352\\ \hline
+Mean & 9.787 & 382.9
+\end{tabular}
+\caption{Indiscriminate vaccinations}
+\end{subfigure}
+\caption{Comparison of 10 runs for different vaccination policies}
+\label{fig:10outs}
+\end{figure}
 
@@
 		\url{http://ithaca.areaconnect.com/statistics.htm}
-	\bibitem{CIC}
+	\bibitem{CIC-stats}
 		\url{http://www.cdc.gov/h1n1flu/surveillanceqa.htm}
+	\bibitem{CIC-qa}
+		\url{http://www.cdc.gov/h1n1flu/qa.htm}
 	\bibitem{walph}
@@
 \section{Source Code}
-	//TODO
+	\lstinputlisting{models/stochastic/spread.m}
+	\lstinputlisting{models/stochastic/random_vaccine.m}
+	\lstinputlisting{models/deterministic/numerical_sir/sir_04.m}
+	\lstinputlisting{models/deterministic/numerical_sir/sir_update_04.m}
+	\lstinputlisting{models/deterministic/sir-two-coupled.m}
+	\lstinputlisting{models/deterministic/logistic-newton-s-investigation.m}
+	\lstinputlisting{models/deterministic/logistic-newton-s-investigation.m}
+
 \section{Individual Contributions}

### 2015-10-01 final_submission/models/deterministic/compute-with-sir.m

@@ -1,0 +1,53 @@
+pkg load odepkg
+
+% contact rate
+global beta = 1.0;
+
+% removal rate
+global gamma = 0.2;
+
+% total population
+global N = 60000;
+
+% vaccination rate
+global v = 4000;
+
+function derivs = sir(t, x)
+	global beta; global gamma; global N; global v;
+	s = x(1); i = x(2); r = x(3);
+	if s >= 0.0
+		derivs = [ - beta * s * i / N - v; beta * s * i / N - gamma * i; gamma * i + v ];
+	else
+		derivs = [ 0; - gamma * I; gamma * I ];
+	end
+end
+
+function [ value, isterminal, parity] = check_for_end(t, x)
+	value = x(1);
+	isterminal = 1;
+	parity = -1;
+end
+
+global options = odeset("RelTol", 1e-10, 
+                 "AbsTol", 1e-8, 
+				 "InitialStep", 0.01, 
+				 "MaxStep", 0.3,
+				 "Events", @check_for_end);
+
+function total_infected = model_with_sir(beta_input, gamma_input)
+	global options;
+	global beta = beta_input;
+	global gamma = gamma_input;
+	global N;
+	global v;
+	[T, Y] = ode45(@sir, [0, 15], [ 0.98; 0.02; 0.0 ]*N, options);
+	
+	total_infected = Y(end, 2) + Y(end, 3) - v*T(end);
+end
+
+%	tx = ty = linspace(0.1, 2.0, 10)';
+%	[betas, gammas] = meshgrid(tx, ty);
+%	mesh( betas, gammas, model_with_sir(betas, gammas));
+
+model_with_sir(1.0, 0.3);
+pause

### 2015-10-01 final_submission/models/deterministic/logistic-newton-s-investigation.m

@@ -1,0 +1,41 @@
+global r = 1.0;
+global K = 1.0;
+global r_v = 0.2;
+global s_0 = 0.05;
+
+function s = s(t)
+	global r; global K; global r_v; global s_0;
+	s = exp(r.*t.*(K - r_v .* t ./ 2)) ./ (1 ./ s_0 + sqrt(pi .* r ./ (2 * r_v)) .* exp( 0.5 * r ./ r_v .* K.^2) .* ( erf( sqrt(0.5 * r ./ r_v) .* (r_v .* t - K)) - erf(- sqrt(0.5 * r ./ r_v) .* K)));
+end
+
+function [s, jac] = minimizee(t)
+	global r; global K; global r_v; global s_0;
+	s = exp(r.*t.*(K - r_v .* t ./ 2)) ./ (1 ./ s_0 + sqrt(pi .* r ./ (2 * r_v)) .* exp( 0.5 * r ./ r_v .* K.^2) .* ( erf( sqrt(0.5 * r ./ r_v) .* (r_v .* t - K)) - erf(- sqrt(0.5 * r ./ r_v) .* K))) - (K - r_v * t);
+	jac = r * s * (K - s - r_v * t) + r_v;
+end
+
+D = linspace(0.01, 0.1, 500);
+R = linspace(0.2, 2.0, 10);
+points = zeros(length(D),1+length(R));
+points(:,1) = D;
+t_1_previous = 8.0;
+for i = 1:length(D)
+	for j = 1:length(R)
+		r = R(j);
+		r_v = D(i) * r_v;
+		t_1 = fsolve(@minimizee, t_1_previous);
+		t_1_previous = t_1;
+		s_1 = s(t_1);
+		r_v = 0.2;
+		points(i,j+1) = s_1;
+	end
+end
+
+csvwrite("data/s-values.csv", points)
+hold on
+for j = 1:length(R)
+	plot(points(:,1), points(:,1+j));
+end
+hold off
+pause
+

### 2015-10-01 final_submission/models/deterministic/logistic-newton-s-plotting-strange.m

@@ -1,0 +1,26 @@
+global r = 1.0;
+global K = 1.0;
+global r_v = 0.2;
+global s_0 = 0.05;
+
+function s = s(t)
+	global r; global K; global r_v; global s_0;
+	s = exp(r.*t.*(K - r_v .* t ./ 2)) ./ (1 ./ s_0 + sqrt(pi .* r ./ (2 * r_v)) .* exp( 0.5 * r ./ r_v .* K.^2) .* ( erf( sqrt(0.5 * r ./ r_v) .* (r_v .* t - K)) - erf(- sqrt(0.5 * r ./ r_v) .* K)));
+end
+
+function [s, jac] = minimizee(t)
+	global r; global K; global r_v; global s_0;
+	s = exp(r.*t.*(K - r_v .* t ./ 2)) ./ (1 ./ s_0 + sqrt(pi .* r ./ (2 * r_v)) .* exp( 0.5 * r ./ r_v .* K.^2) .* ( erf( sqrt(0.5 * r ./ r_v) .* (r_v .* t - K)) - erf(- sqrt(0.5 * r ./ r_v) .* K))) - (K - r_v * t);
+	jac = r * s * (K - s - r_v * t) + r_v;
+end
+
+D = linspace(0.06, 1.0, 10);
+T = linspace(0, 20, 1000);
+hold on
+for i = 1:length(D)
+	r_v = 0.2*D(i);
+	plot(T, s(T));
+end
+hold off
+pause
+

### 2015-10-01 final_submission/models/deterministic/numerical_sir/sir_01.m

@@ -1,0 +1,28 @@
+clc; close all; clear;
+pkg load odepkg
+%===========%
+% SIR model %
+%===========%
+
+% ----- Model Parameters ----- %
+s0 = .99;
+i0 = .01;
+r0 = .0;
+
+d = .25;
+nu = 1/d;
+R0 = 1.5;
+beta = R0 * nu;
+% ---------------------------- %
+
+t0 = 0;
+tf = 8;
+y0 = [s0;i0;r0];
+[T,Y] = ode45(@(t,y) sir_update_01(y, beta, nu), [t0, tf], y0);
+
+figure
+plot(T,Y)
+xlabel('Months since outbreak')
+ylabel('Fraction of population')
+legend('Susceptible', 'Infected', 'Removed')
+pause

### 2015-10-01 final_submission/models/deterministic/numerical_sir/sir_03.m

@@ -1,0 +1,36 @@
+clc; close all; clear;
+%===================================%
+% SIR model with death and vaccines %
+%===================================%
+
+% ----- Model Parameters ----- %
+s0 = .95;
+i0 = .05;
+r0 = .0;
+d0 = .0;
+N = 60000;
+
+d = .25;
+nu = 1/d;
+R0 = 1.5;
+beta = R0 * nu;
+mu = 0.00001709; % * 20;
+vaccines = 4000;
+gamma = vaccines/N;
+% ---------------------------- %
+
+t0 = 0;
+tf = 8;
+y0 = [s0;i0;r0;d0];
+[T,Y] = ode45(@(t,y) sir_update_03(y, beta, nu, mu, gamma), [t0, tf], y0);
+ 
+figure
+plot(T,Y(:,1:3))
+ylim([0,1])
+xlabel('Months since outbreak')
+ylabel('Fraction of population')
+legend('Susceptible', 'Infected', 'Removed')
+
+epidemic_duration = T(find(Y(:,2) < .01 * max(Y(:,2)),1)) % months
+total_deaths = Y(end,4) * N                               % people
+total_suffering = trapz(T,Y(:,2)) * N                     % person-months

### 2015-10-01 final_submission/models/deterministic/numerical_sir/sir_update_02.m

@@ -1,0 +1,10 @@
+function [y_prime] = sir_update_02(y,beta,nu,mu)
+s = y(1); 
+i = y(2);
+y_prime = zeros(3,1);
+y_prime(1) = -beta * s * i;
+y_prime(2) = i * (beta * s - nu - mu);
+y_prime(3) = nu * i;
+y_prime(4) = mu * i;
+end
+

### 2015-10-01 final_submission/models/deterministic/numerical_sir/sir_update_c.m

@@ -1,0 +1,13 @@
+function [y_prime] = sir_update_04(y,beta,nu,mu,gamma)
+s = y(1:3);
+i = y(4:6);
+
+y_prime = zeros(9,1);
+for k = 1:3
+    y_prime(k) = -beta(k,:) * i * s(k) - gamma(k);
+    y_prime(k+3) = beta(k,:) * i * s(k) - i(k) * (nu + mu(k));
+    y_prime(k+9) = nu * i(k) + gamma(k);
+    y_prime(k+12) = mu(k) * i(k);
+end
+
+end

### 2015-10-01 final_submission/models/deterministic/sir.m

@@ -1,0 +1,55 @@
+pkg load odepkg
+
+% contact rate
+global beta;
+
+% removal rate
+global gamma;
+
+% total population
+global N = 60000;
+
+% vaccination rate
+global v = 4000;
+
+function derivs = sir(t, x)
+	global beta; global gamma; global N; global v;
+	s = x(1); i = x(2); r = x(3);
+	derivs = [ - beta * s * i / N - v; beta * s * i / N - gamma * i; gamma * i + v ];
+end
+
+function [ value, isterminal, parity] = check_for_end(t, x)
+	value = x(1);
+	isterminal = 1;
+	parity = -1;
+end
+
+options = odeset("RelTol", 1e-2, 
+                 "AbsTol", 1e-2, 
+				 "InitialStep", 0.01, 
+				 "MaxStep", 0.1,
+				 "Events", @check_for_end)
+
+warning("off", "all")
+Beta = linspace(0.6, 2.0, 20);
+Gamma = linspace(0.1, 0.45, 20);
+results = zeros(length(Beta), length(Gamma));
+for i = 1:length(Beta)
+	for j = 1:length(Gamma)
+		beta = Beta(i);
+		gamma = Gamma(j);
+		beta
+		gamma
+		[T, Y] = ode45(@sir, [0, 15.0], [ 0.98; 0.02; 0.0 ]*N, options);
+		results(i,j) = Y(end, 2) + Y(end, 3) - v*T(end);
+		Y(end, 2) + Y(end, 3) - v*T(end)
+	end
+end
+
+[bmesh, gmesh] = meshgrid(Beta, Gamma);
+mesh(bmesh, gmesh, results)
+csvwrite('../../data/sir-parameter.csv', results);
+results
+%plot(T, Y)
+%legend('Seceptible','Infected','Removed')
+pause

### 2015-10-01 final_submission/models/stochastic/spread.m

@@ -1,0 +1,156 @@
+function [peak, hosps, sick_per_delta, hospitalized] = spread(age_range, p)
+  if nargin == 1
+    p = false;
+  end
+
+  global ithaca_pop;
+  global vaccines;
+  global init_sick;
+  global num_juniors;
+  global num_adults;
+  global num_seniors;
+  global junior_center;
+  global adult_center;
+  global senior_center;
+
+  ithaca_pop = 60000;
+  vaccines = 4000/4; % Vaccines per week
+
+  ithaca_pop = ithaca_pop;
+  vaccines = vaccines;
+
+  init_sick = ithaca_pop/20;
+
+  % Census data for age ranges
+  num_juniors   = 0.6*ithaca_pop;
+  num_adults    = 0.3*ithaca_pop;
+  num_seniors   = 0.1*ithaca_pop;
+
+  junior_center   = num_juniors/2;
+  adult_center    = num_juniors + num_adults/2;
+  senior_center   = num_juniors + num_adults + num_seniors/2;
+
+  heal_chance = 0.9;
+  age = zeros(1, ithaca_pop);
+  sick = zeros(1, ithaca_pop);
+  hospitalized = zeros(1, ithaca_pop);
+  vaccinated = zeros(1, ithaca_pop);
+
+  function v = ifelse(c, t, f)
+    if c
+      v = t;
+    else
+      v = f;
+    end
+  end
+
+  function conn = connectivity(age)
+    switch age
+      case 1
+        conn = 25;
+      case 2
+        conn = 15;
+      case 3
+        conn = 5;
+    end
+  end
+
+  function comp = comp_chance(age)
+    switch age
+      case 1
+        comp = 0.05;
+      case 2
+        comp = 0.2;
+      case 3
+        comp = 0.75;
+    end
+    % comp = 0.95;
+  end
+
+  for i = 1:ithaca_pop
+    if i <= num_juniors
+      age(i) = 1;
+    elseif i <= (num_juniors + num_adults)
+      age(i) = 2;
+    else
+      age(i) = 3;
+    end
+  end
+
+  weeks = 8; % 2 months
+  sick_per_delta = zeros(1 + weeks, 4);
+
+  for c = round(rand(1,init_sick)*(ithaca_pop - 1)) + 1
+    sick(c) = true;
+    sick_per_delta(1, age(c) + 1) = sick_per_delta(1, age(c) + 1) + 1;
+  end
+
+  n = 1;
+  for w = 1:weeks
[…]

### 2015-10-01 final_submission/manuscript.tex
figures in

--- 2015-10-01 final_submission/manuscript.tex
@@
 \subsection{Predictions}
-	//TODO
-% \subsection{Validation and Robustness Testing}
+	Running the $sir_{04}.m$ script to solve the system of ODE's for the segmented SIR model with equal distribution of vaccines produces the following plots:
+
+    \begin{figure}[H]
+    \centering
+        \includegraphics[width=\textwidth]{figures/ryan_plot_01.png}
+    \end{figure}
+    \begin{figure}[H]
+    \centering
+        \includegraphics[width=\textwidth]{figures/ryan_plot_02.png}
+    \end{figure}
+    \begin{figure}[H]
+    \centering
+        \includegraphics[width=\textwidth]{figures/ryan_plot_03.png}
+    \end{figure}
+    \begin{figure}[H]
+    \centering
+        \includegraphics[width=\textwidth]{figures/ryan_plot_04.png}
+    \end{figure}
+
+    The fraction of deaths remains small for all populations so it is not plotted. The total number of person-months of suffering is the integral of the green curve.
+% \subsection{Analysis of Model (robustness)}
 % 	//TODO

### 2015-10-01 final_submission/manuscript.tex
Added contributions

@@
 \section{Individual Contributions}
-	//TODO
+	Paul handled mostly what covered the Stochastic Model, while Chris and Ryan worked on their own deterministic models.
+Everyone did a lot of their work in the presence of everyone else, while still putting in a lot of time at home.
+When it came time to submit, everyone was ready to work.
+All in all, lots of work put in from all team member and we were all satisfied with each other's work.
 

### 2015-10-01 final_submission/manuscript.tex
done SIR (mostly)

--- 2015-10-01 final_submission/manuscript.tex
@@
     The fraction of deaths remains small for all populations so it is not plotted. The total number of person-months of suffering is the integral of the green curve.
+
+    Running this with vaccine distributions (.1, .4, .4, .1) results in the following MATLAB output:
+
+    \begin{lstlisting}
+        total_deaths =
+
+            0.0030    0.2532    0.0529    0.0049
+
+        total_suffering =
+
+            1.0e+03 *
+
+            0.0389    7.0540    3.0925    0.1686
+    \end{lstlisting}
+
+    Running this with vaccine distributions (.4, .1, .1, .4) results in the following MATLAB output:
+
+    \begin{lstlisting}
+        total_deaths =
+
+            0.0020    0.2674    0.0579    0.0039
+
+        total_suffering =
+
+            1.0e+03 *
+
+            0.0263    7.4504    3.3864    0.1356
+    \end{lstlisting}
+    Running this with vaccine distributions (.1, .4, .4, .1) and $\mu = \mu * 20$results in the following MATLAB output:
+
+     \begin{lstlisting}
+        total_deaths =
+
+            0.0598    5.0633    1.0570    0.0979
+
+        total_suffering =
+
+            1.0e+03 *
+
+            0.0389    7.0522    3.0919    0.1685
+    \end{lstlisting}
+
+    Running this with vaccine distributions (.4, .1, .1, .4) and $\mu = \mu * 20$results in the following MATLAB output:
+
+    \begin{lstlisting}
+        total_deaths =
+
+            0.0403    5.3464    1.1572    0.0787
+
+        total_suffering =
+
+            1.0e+03 *
+
+            0.0263    7.4485    3.3857    0.1356
+    \end{lstlisting}
 % \subsection{Analysis of Model (robustness)}

### 2015-10-01 final_submission/manuscript.tex
Some touchups

--- 2015-10-01 final_submission/manuscript.tex
@@
 
-Which unfortunately doesn't look promising.
+Which unfortunately doesn't look promising. When the disease is twenty times as deadly, hospitalization rates scale up accordingly, and vaccination policies keep their effectiveness.
 
@@
 We see that the random policy actually puts just about as many people in the hospital as the policy that targets seniors, while keeping the peak infection rate actually lower than all the policies.
+When the disease is deadlier, all models agree that the random policy is better than all.
+\par Consensus on how long a given epidemic should last is about 2 months, and it seems that our model agrees, so we can at least verify that it decently approximates a real spread.
+\par Part of the reason why the different methods don't have such a different outcome seems to be that the disease simply spreads too fast.
+Long before enough vaccinations can be handed out, the infection has spread too far for it to be slowed down.
 

### 2015-10-01 final_submission/manuscript.tex
I'm done

--- 2015-10-01 final_submission/manuscript.tex
@@
     \end{lstlisting}
-    Running this with vaccine distributions (.1, .4, .4, .1) and $\mu = \mu * 20$results in the following MATLAB output:
+    Running this with vaccine distributions (.1, .4, .4, .1) and $\mu = \mu * 20$ results in the following MATLAB output:
 
@@
 
-    Running this with vaccine distributions (.4, .1, .1, .4) and $\mu = \mu * 20$results in the following MATLAB output:
+    Running this with vaccine distributions (.4, .1, .1, .4) and $\mu = \mu * 20$ results in the following MATLAB output:
 
@@
     \end{lstlisting}
-% \subsection{Analysis of Model (robustness)}
-% 	//TODO
+    \subsection{Analysis of Model}
+	 This model suffer from being dependent upon too many parameters, and it is hard to come up with an optimal vaccine distribution.
+Tweaking the parameters a little bit can have a drastic effect on the model, and it relies too much on our research being accurate.
+In the future I would search for a pareto optimal distribution.
+Also, I would fix the assumption of a linear vaccination rate so we are not wasting vaccines on a completely vaccinated population, and can test getting all infants out of the way at the beginning.
 % \subsection{Strengths and Weaknesses}

### 2015-10-01 final_submission/manuscript.tex
Final commit. DONE BITCHES YEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEEE

--- 2015-10-01 final_submission/manuscript.tex
@@
 
-\section{Conclusion}
-	//TODO
+\section{Conclusions}
+
+A letter to the Ithaca Journal from the Modeling Department of the CDC:
+
+\paragraph{}
+This letter recommends a first-come first-serve
+	vaccination policy in Ithaca during an epidemic.
+My colleagues and I are epidemic modelers, and we have researched
+	past flu outbreaks and used them to simulate how Ithaca
+	would react to a flu outbreak.
+Based on our research and modeling, we determined that if a
+	flu epidemic strikes Ithaca suddenly, it is more important
+	for all of the vaccines to be distributed than for us to
+	vaccinate a particular population first.
+We urge people to vaccinate before the outbreak, so that they are not
+	vulnerable to the disease at the beginning.
+By not vaccinating yourself, you not only risk catching the flu,
+	you risk spreading it to others, possibly young children and the
+	elderly who are less able to resist it.
+We urge Cornell, Ithaca College, and the Tompkins County Public School
+	System, to provide vaccination in class, especially before an epidemic.
+
+It is important to vaccinate before an outbreak because vaccines are
+	distributed to each reason on a monthly basis.
+We must not waste our distributed portion of vaccines before the outbreak
+	starts.
+We recieve 4000 every month, and we want to make sure that people are
+	ready for the flu season before it starts.
+Therefore, everybody should be vaccinated, especially children and the elderly.
+
+During an outbreak, we recommend that vaccines should be uniformly distributed
+	between Cornell, Ithaca College, and Ithaca Proper.
+Our models did not find a significant difference between distribution strategies,
+	so we urge local officials and school administrators to estimate how
+	many students they think they can vaccinate each month.
+It is important that these valuable vaccinations not go to waste.
+
 \section{References}
